--- 
title_meta  : Chapter 6
title       : Data Frames
description : Most data sets you will be working with will be stored as a data frame. By the end of this chapter, you will be able to create a data frame, select interesting parts of a data frame and order a data frame according to certain variables.
attachments : 
  slides_link: https://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/slides/ch6_slides.pdf

--- type:VideoExercise lang:r xp:50 skills:1 key:d4bde604ab
## Explore the Data Frame

*** =video_link
//player.vimeo.com/video/138173996

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch6_1.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:d4ddcf9a7d
## Have a look at your data set

Working with large data sets is not uncommon in data analysis. When you work with (extremely) large data sets and data frames, your first task as a data analyst is to develop a clear understanding of its structure and main elements. Therefore, it is often useful to show only a small part of the entire data set. 

There are several ways to do this in R. The function [`head()`](http://www.rdocumentation.org/packages/utils/functions/head) enables you to show the first observations of a data frame (or any R object you pass to it). Unoriginally, the function [`tail()`](http://www.rdocumentation.org/packages/utils/functions/head) prints out the last observations in your data set. You can also use the function [`dim()`](http://www.rdocumentation.org/packages/base/functions/dim) to show the dimensions of your data set.

In this exercise, you'll be working with the `mtcars` dataset, that is available in R by default.

*** =instructions
- Print the first observations of the [`mtcars`](http://www.rdocumentation.org/packages/datasets/functions/mtcars) data set.
- Use the [`tail()`](http://www.rdocumentation.org/packages/utils/functions/head) function to display the last observations.
- Finally, display the overall dimensions of the [`mtcars`](http://www.rdocumentation.org/packages/datasets/functions/mtcars) data frame with [`dim()`](http://www.rdocumentation.org/packages/base/functions/dim).

*** =hint
You'll need [`head()`](http://www.rdocumentation.org/packages/utils/functions/head) to show the first observations in [`mtcars`](http://www.rdocumentation.org/packages/datasets/functions/mtcars).

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Print the first observations of mtcars


# Print the last observations of mtcars


# Print the dimensions of mtcars
```

*** =solution
```{r}
# Print the first observations of mtcars
head(mtcars)

# Print the last observations of mtcars
tail(mtcars)

# Print the dimensions of mtcars
dim(mtcars)
```

*** =sct
```{r}
test_error()
test_function("head", "x", incorrect_msg = "Have you specified the correct parameter in the [`head()`](http://www.rdocumentation.org/packages/utils/functions/head) function? Make sure to pass it a data set you want to inspect, `mtcars` in this case.")
test_function("tail", "x", incorrect_msg = "Have you specified the correct parameter in the [`tail()`](http://www.rdocumentation.org/packages/utils/functions/head) function? Make sure to pass it a data set you want to inspect, `mtcars` in this case.")
test_output_contains("dim(mtcars)", incorrect_msg = "Don't forget to also call the [`dim()`](http://www.rdocumentation.org/packages/base/functions/dim) function on `mtcars`!")

success_msg("Wonderful! So, do you now have a good idea about what we have in the data set? For a full overview of the variables' meaning, type `?mtcars` in the console and read the help page. Continue to the next exercise!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:c8f389fdbd
## Have a look at the structure

Another method that is often used to get a rapid overview of your data is the function [`str()`](http://www.rdocumentation.org/packages/utils/functions/str). The function [`str()`](http://www.rdocumentation.org/packages/utils/functions/str) shows you the structure of your data set. For a data frame it tells you:

- The total number of observations (e.g. 32 car types)
- The total number of variables (e.g. 11 car features)
- A full list of the variables names (e.g. mpg, cyl ... )
- The data type of each variable (e.g. num for car features)
- The first observations

Applying the [`str()`](http://www.rdocumentation.org/packages/utils/functions/str) function will often be the first thing that you do when receiving a new data set or data frame. It is a great way to get more insight in your data set before diving into the real analysis.

*** =instructions
Investigate the structure of [`mtcars`](http://www.rdocumentation.org/packages/datasets/functions/mtcars). Make sure that you see the same numbers, variables and data types as mentioned above.

*** =hint
Use the [`str()`](http://www.rdocumentation.org/packages/utils/functions/str) function with [`mtcars`](http://www.rdocumentation.org/packages/datasets/functions/mtcars) as input!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Investigate the structure of the mtcars data set
```

*** =solution
```{r}
# Investigate the structure of the mtcars data set
str(mtcars)
```

*** =sct
```{r}
test_function("str","object",incorrect_msg = "Make sure to check the structure of the `mtcars` data set.")
test_output_contains("str(mtcars)", incorrect_msg = "Make sure that you use the [`str()`](http://www.rdocumentation.org/packages/utils/functions/str) function on `mtcars`.")
success_msg("Nice work! Can you find all the information that is listed in the exercise's assignment? Continue to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:f09c0189ac
## Creating a data frame (1)

Since using built-in data sets is not even half the fun of creating your own data sets, the next exercises are based on your personally developed data set. So put your jet pack on because it is time for some good old fashioned space exploration! 

As a first goal, you want to construct a data frame that describes the main characteristics of eight planets in our solar system. According to your good friend Buzz, the main features of a planet are:

- The type of the planet (Terrestrial or Gas Giant).
- The planet's diameter relative to the diameter of the Earth.
- The planet's rotation across the sun relative to that of the Earth.
- If the planet has rings or not (TRUE or FALSE).

After doing some high-quality research on [Wikipedia](http://en.wikipedia.org/wiki/Planet), you feel confident enough to create the necessary vectors: `planets`, `type`, `diameter`, `rotation` and `rings`. Can you correctly use the [`data.frame()`](http://www.rdocumentation.org/packages/base/functions/data.frame) function to create a data set from this information?

*** =instructions
- Use the function [`data.frame()`](http://www.rdocumentation.org/packages/base/functions/data.frame) to construct `planets_df`.
- Make sure that you've actually created a data frame with 8 observations and 5 variables with [`str()`](http://www.rdocumentation.org/packages/utils/functions/str).

*** =hint
The [`data.frame()`](http://www.rdocumentation.org/packages/base/functions/data.frame) function takes as arguments the vectors that will become the columns of the data frame, separated by commas. The columns in this case are (in this order): `planet`, `type`, `diameter`, `rotation` and `rings`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Definition of vectors
planets <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create a data frame: planets_df


# Display the structure of planets_df

```

*** =solution
```{r}
# Definition of vectors
planets <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create a data frame: planets_df
planets_df <- data.frame(planets, type, diameter, rotation, rings)

# Display the structure of planets_df
str(planets_df)
```

*** =sct
```{r}
test_correct({
  test_object("planets_df",
              undefined_msg = "Please make sure to define a variable `planets_df`.",
              incorrect_msg = "Make sure to assign the correct order of arguments to the data.frame `planets_df`. The correct order is planets, type, diameter, rotation and rings.")  
}, {
  msg = "Do not change anything about the definition of the vector. Only add code to create the `planets_df` data frame!"
  test_object("planets", undefined_msg = msg, incorrect_msg = msg)
  test_object("type", undefined_msg = msg, incorrect_msg = msg)
  test_object("diameter", undefined_msg = msg, incorrect_msg = msg)
  test_object("rotation", undefined_msg = msg, incorrect_msg = msg)
  test_object("rings", undefined_msg = msg, incorrect_msg = msg)
})

test_output_contains("str(planets_df)", incorrect_msg = "Don't forget to display the structure of `planets_df`!")

success_msg("Great job! The structure of `planets_df` reveals that both the `planets` as the `type` column are factors, and not character vectors. That's not really what you want, right?")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:7090dc3538
## Creating a data frame (2)

In the previous exercise, you found out that both the `planets` and `type` columns of `planets_df` are factors. For the `type` column this makes sense, because a planet type is some kind of category. For the `planets` column, however, that contains the planet names, this is less logical.

You can set the `stringsAsFactors` argument inside [`data.frame()`](http://www.rdocumentation.org/packages/base/functions/data.frame) to avoid that R automatically converts character vectors to factors:

```
data.frame(vec1, vec2, ..., stringsAsFactors = FALSE)
```

Up to you now to adapt the way `planets_df` is constructed!

*** =instructions
- Encode the `type` vector in a factor, called `type_factor`.
- Next use `planets`, `type_factor`, `diameter`, `rotation` and `rings` to construct `planets_df`. This time, make sure that strings are not converted to factors, by setting `stringsAsFactors = FALSE`.
- Display the structure of `planets_df` to check you coded things correctly.

*** =hint
Use the function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) to encode `type` as a factor.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Definition of vectors
planets <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet",
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant","Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Encode type as a factor: type_factor


# Construct planets_df: strings are not converted to factors!


# Display the structure of planets_df

```

*** =solution
```{r}
# Definition of vectors
planets <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Encode type as a factor: type_factor
type_factor <- factor(type)

# Construct planets_df: strings are not converted to factors!
planets_df <- data.frame(planets, type_factor, diameter, rotation, rings, stringsAsFactors = FALSE)

# Display the structure of planets_df
str(planets_df)
```

*** =sct
```{r}
msg <- "Do not remove or change the definition of all the vectors!"
test_object("planets", undefined_msg = msg, incorrect_msg = msg)
test_object("type", undefined_msg = msg, incorrect_msg = msg)
test_object("diameter", undefined_msg = msg, incorrect_msg = msg)
test_object("rotation", undefined_msg = msg, incorrect_msg = msg)
test_object("rings", undefined_msg = msg, incorrect_msg = msg)

test_object("type_factor", incorrect_msg = "Have you correctly created `type_factor`? Simply use the [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) function on `type`.")
test_object("planets_df", incorrect_msg = "Have you correctly created `planets_df`? Make sure to use `type_factor` instead of `type` and set `stringsAsFactors` to `FALSE` inside [`data.frame()`](http://www.rdocumentation.org/packages/base/functions/data.frame).")
test_output_contains("str(planets_df)", incorrect_msg = "Don't forget to display the structure of `planets_df`.")

success_msg("That looks more like it! Head over to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:a80ae7fbe8
## Rename the data frame columns

As a data frame is actually a list containing same-length vectors under the hood, it's possible to name and rename data frames just as you did with lists. To create a data frame and name it in one and the same call you can use:

```
data.frame(name1 = vec1, name2 = vec2, ...)
```

You can also name a data frame after creating it:

```
my_df <- data.frame(vec1, vec2, ...)
names(my_df) <- c("name1", "name2", ...)
```

Very proud of your first ever data frame, you show it to your friend Buzz. He's pretty impressed that you managed to include both factor and character columns, but he still finds the column names pretty odd. Time to make some improvements! The code that constructs the improved data frame, as you coded in the previous exercise, is already included.

*** =instructions
Rename the columns of `planets_df`. As `planets_df` is already created, you'll want to use the [`names()`](http://www.rdocumentation.org/packages/base/functions/names) function. 

- Name the `planets` column `name`.
- Name the `type_factor` column `type`.
- You can keep the names `diameter` and `rotation`.
- Change the name `rings` to `has_rings`.

Finally, print `planets_df` after you renamed it (not its structure!).

*** =hint
You'll need the vector containing `"name"`, `"type"`, `"diameter"`, `"rotation"` and `"has_rings"`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Construct improved planets_df
planets <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)
type_factor <- factor(type)
planets_df <- data.frame(planets, type_factor, diameter, rotation, rings, stringsAsFactors = FALSE)

# Improve the names of planets_df


# Print planets_df

```

*** =solution
```{r}
# Construct improved planets_df
planets <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)
type_factor <- factor(type)
planets_df <- data.frame(planets, type_factor, diameter, rotation, rings, stringsAsFactors = FALSE)

# Improve the names of planets_df
names(planets_df) <- c("name", "type", "diameter", "rotation", "has_rings")

# Print planets_df
planets_df
```

*** =sct
```{r}

msg <- "Do not remove or change the definition of the predefined variables!"
test_object("planets", undefined_msg = msg, incorrect_msg = msg)
test_object("type", undefined_msg = msg, incorrect_msg = msg)
test_object("diameter", undefined_msg = msg, incorrect_msg = msg)
test_object("rotation", undefined_msg = msg, incorrect_msg = msg)
test_object("rings", undefined_msg = msg, incorrect_msg = msg)
test_object("type_factor", undefined_msg = msg, incorrect_msg = msg)
test_object("planets_df", undefined_msg = msg, incorrect_msg = "Don't change the contents of `planets_df`, only change the column names!")
test_object("planets_df", eq_condition = "equal",
            undefined_msg = msg, incorrect_msg = "Are you sure you have correctly renamed the columns of `planets_df`? The hint might be able to help you out.")

test_output_contains("planets_df", incorrect_msg = "Don't forget to print `planets_df`.")
success_msg("Nice one! This is going fast!")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:9a2f941de8
## Subset, Extend & Sort Data Frames

*** =video_link
//player.vimeo.com/video/138174008

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch6_2.master.m3u8


--- type:NormalExercise lang:r xp:100 skills:1 key:b6125af738
## Selection of data frame elements

Similar to matrices, you select elements from a data frame with the help of square brackets `[ ]`. By using a comma, you can indicate what to select from the rows and the columns respectively:

```
# first row, second column
my_df[1,2]

# rows 1 to 3, columns 2 to 4
my_df[1:3,2:4]

# Entire first row
my_df[1, ]

# rows 1 to 3 of "type" column
planets_df[1:3,2]
planets_df[1:3,"type"]
```

Let us now apply this technique on `planets_df`! This data frame is already available in the workspace.

*** =instructions
- Select the type of Mars; store the factor in `mars_type`.
- Store the entire rotation column in `rotation` as a vector.
- Create a data frame, `closest_planets_df`, that contains all data on the first three planets.
- Likewise, build the data frame `furthest_planets_df` that contains all data on the last three planets.

*** =hint
`planets_df[1:3,]` will select all elements of the first three rows.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter6.RData"))
```

*** =sample_code
```{r}
# planets_df is pre-loaded

# The type of Mars: mars_type


# Entire rotation column: rotation


# First three planets: closest_planets_df


# Last three planets: furthest_planets_df


```

*** =solution
```{r}
# planets_df is pre-loaded

# The type of Mars: mars_type
mars_type <- planets_df[4, 2]

# Entire rotation column: rotation
rotation <- planets_df[ ,4]

# First three planets: closest_planets_df
closest_planets_df <- planets_df[1:3, ]

# Last three planets: furthest_planets_df
furthest_planets_df <- planets_df[6:8, ]
```

*** =sct
```{r}

msg <- "Do not remove or overwrite the `planets_df` data frame!"
test_object("planets_df", undefined_msg = msg, incorrect_msg = msg)

test_object("mars_type", 
            incorrect_msg = "Are you sure you correctly selected the type of Mars?")
test_object("rotation", 
            incorrect_msg = "Have another look at the command to define `rotation`. You'll want to select the fourth column.")
test_object("closest_planets_df", 
            incorrect_msg = "Did you select the first three rows of `planets_df` to create `closest_planets_df`?")
test_object("furthest_planets_df",
            incorrect_msg = "Make sure that you select the last three rows of `planets_df` to build `furthest_planets_df`.")

success_msg("Great! Feel free to have a look at the variables you've just created. Apart from selecting elements from your data frame by index, you can also use the column names.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:7aa0a94261
## Only planets with rings (1)

You will often want to select an entire column, namely one specific variable from a data frame. If you want to select the column `diameter` from `planets_df`, you can use either on of the following:

```
planets_df[, 3]
planets_df[, "diameter"]
planets_df$diameter
```

*** =instructions
- Make use of the `$` sign to create the variable `rings_vector` that contains the entire `has_rings` column in the `planets_df` data frame.
- Print the `rings_vector`; it should be a vector.

*** =hint
`my_df$col_name` is the most convenient way to select a column from a data frame. In this case, the data frame is `planets_df` and the variable is `has_rings`.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter6.RData"))
```

*** =sample_code
```{r}
# planets_df is pre-loaded in your workspace

# Create rings_vector


# Print rings_vector

```

*** =solution
```{r}
# planets_df is pre-loaded in your workspace

# Create rings_vector
rings_vector <- planets_df$has_rings

# Print rings_vector
rings_vector
```

*** =sct
```{r}
msg <- "Do not remove or overwrite the `planets_df` data frame!"
test_object("planets_df", undefined_msg = msg, incorrect_msg = msg)

test_object("rings_vector", incorrect_msg = "It looks like `rings_vector` does not contain all the elements of the `has_rings` variable of`planets_df`.")

test_output_contains("rings_vector", incorrect_msg = "Don't forget to print `rings_vector`!")

success_msg("Great! Continue to the next exercise and discover yet another way of subsetting!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:d6245c1bb1
## Only planets with rings (2)

You probably remember from high school that some planets in our solar system have rings and others do not. But due to other priorities at that time (read: puberty) you can not recall their names, let alone their rotation speed, etc. Could R help you out?

The `rings_vector` you've coded before is a logical vector. It's `TRUE` when the corresponding planets have rings and `FALSE` when they don't. To select those observations from `planets_df` that have rings, you can use the `rings_vector` and perform subsetting by logicals!

To subset observations by logicals, put the logical vector and a comma inside square brackets, similar to this:

```
df[,logical_vector]
```

*** =instructions
- Assign to `planets_with_rings_df` all data in the `planets_df` data set for the planets with rings, that is, where `rings_vector` is `TRUE`.
- Print `planets_with_rings_df`.

*** =hint
Select elements from `planets_df` by using the square brackets. The `rings_vector` contains boolean values and R will select only those rows/columns were the vector element is `TRUE`. In this case, you want to select rows based on `rings_vector` and select all the columns.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter6.RData"))
```

*** =sample_code
```{r}
# planets_df pre-loaded in your workspace

# Create rings_vector
rings_vector <- planets_df$has_rings

# Select the information on planets with rings: planets_with_rings_df


# Print planets_with_rings_df
```

*** =solution
```{r}
# planets_df pre-loaded in your workspace

# Create rings_vector
rings_vector <- planets_df$has_rings

# Select the information on planets with rings: planets_with_rings_df
planets_with_rings_df <- planets_df[rings_vector,]

# Print planets_with_rings_df
planets_with_rings_df
```

*** =sct
```{r}

msg <- "Do not remove or overwrite `planets_df` or `rings_vector`!"
test_object("planets_df", undefined_msg = msg, incorrect_msg = msg)
test_object("rings_vector", undefined_msg = msg, incorrect_msg = msg)

test_object("planets_with_rings_df",
            incorrect_msg = "It looks like `planets_with_rings_df` does not contain all the data of the planets with rings. Make sure to not specify any column selector, to keep all columns.")

test_output_contains("planets_with_rings_df",
                     incorrect_msg = "Don't forget to print `planets_with_rings_df`!")
success_msg("Nice work, but this is a rather tedious solution. The next exercise will teach you how to do it in a more concise way.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:c1a08e245c
## Only planets with rings but shorter

So what exactly did you learn in the previous exercises? You selected a subset from a data frame (`planets_df`) based on whether or not a certain condition was true (rings or no rings), and you managed to pull out all relevant data. Pretty awesome! By now, NASA is probably already flirting with your CV!

Instead of having to define a vector `rings_vector`, which you then use to subset `planets_df`, you could've also used either one of these:

```
planets_df[planets_df$has_rings, ]
planets_df[planets_df$has_rings == TRUE, ]
```

*** =instructions
- Create a data frame `small_planets_df` with planets that have a diameter smaller than the Earth. This means that the `diameter` variable should be smaller than 1, since diameter is a relative measure of the planet's diameter in relation to planet Earth.
- Build another data frame, `slow_planets_df`, with the observations that have a longer rotation period than Earth. This means that the absolute value (use the function [`abs()`](http://www.rdocumentation.org/packages/base/functions/MathFun)) of the `rotation` variable should be greater than 1.

*** =hint
Make use of the logical operators `>` and `<`. Use the [`abs()`](http://www.rdocumentation.org/packages/base/functions/MathFun) function for absolute values. 

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter6.RData"))
```

*** =sample_code
```{r}
# planets_df is pre-loaded in your workspace

# Planets that are smaller than planet Earth: small_planets_df


# Planets that rotate slower than planet Earth: slow_planets_df

```

*** =solution
```{r}
# planets_df is pre-loaded in your workspace

# Planets that are smaller than planet Earth: small_planets_df
small_planets_df <- planets_df[planets_df$diameter < 1, ]        # option 1
small_planets_df  <- subset(planets_df, subset = diameter < 1)   # option 2

# Planets that rotate slower than planet Earth: slow_planets_df
slow_planets_df <- planets_df[abs(planets_df$rotation) > 1, ]      # option 1
slow_planets_df <- subset(planets_df, subset = abs(rotation) > 1)  # option 2
```

*** =sct
```{r}

msg <- "Do not remove or overwrite the `planets_df` data frame!"
test_object("planets_df", undefined_msg = msg, incorrect_msg = msg)

test_object("small_planets_df",
            incorrect_msg = "It looks like `small_planets_df` does not contain the correct subset of `planets_df`.")

test_object("slow_planets_df",
            incorrect_msg = "It looks like `slow_planets_df` does not contains the correct subset of `planets_df`. Make sure to use the [`abs()`](http://www.rdocumentation.org/packages/base/functions/MathFun) function for absolute values.")

success_msg("Great! Not only is the [`subset()`](http://www.rdocumentation.org/packages/base/functions/subset) function more concise, it is probably also more understandable for people who read your code. Continue to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:e9ca3eeb99
## Add variable/column

There are many cases in which you'll want to add more variables to your data frame. This comes down to adding a column to the data frame. The exact same techniques to select columns from a data frame can be used here. To add a column `new_column` to `my_df`, with data from `my_vec`, you can use one of the following calls:

```
my_df$new_column <- my_vec
my_df[["new_column"]] <- my_vec
my_df <- cbind(my_df, new_column = my_vec)
```

You've browsed [Wikipedia](https://en.wikipedia.org/wiki/Planet) and also decide to add a column that lists the number of moons each of the planets has. Also the planets' masses can be a cool addition. The `moon` and `masses` vectors are already included in the workspace; up to you to add them to `planets_df`.

*** =instructions
- Add `moons` to `planets_df` under the variable name "moon".
- In a similar fashion, add `masses` under the variable name `"mass"`.

*** =hint
To add a new column called "moon", you can use: `planets_df$moon <- moons`.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter6.RData"))
```

*** =sample_code
```{r}
# planets_df is already pre-loaded in your workspace

# Definition of moons and masses
moons <- c(0, 0, 1, 2, 67, 62, 27, 14)
masses <- c(0.06, 0.82, 1.00, 0.11, 317.8, 95.2, 14.6, 17.2)

# Add moons to planets_df under the name "moon"


# Add masses to planets_df under the name "mass"

```

*** =solution
```{r}
# planets_df is already pre-loaded in your workspace

# Definition of moons and masses
moons <- c(0, 0, 1, 2, 67, 62, 27, 14)
masses <- c(0.06, 0.82, 1.00, 0.11, 317.8, 95.2, 14.6, 17.2)

# Add moons to planets_df under the name "moon"
planets_df$moon <- moons

# Add masses to planets_df under the name "mass"
planets_df$mass <- masses
```

*** =sct
```{r}

undef_msg <- "Do not remove `planets_df`!"
msg <- "Do not change anything about the columns that were already in `planets_df`; you should only <i>add</i> columns."
test_data_frame(name = "planets_df", 
                columns = c("name", "type", "diameter", "rotation", "has_rings"),
                undefined_msg = undef_msg, undefined_cols_msg = msg, incorrect_msg = msg)

test_data_frame(name = "planets_df",
                columns = "moon",
                undefined_cols_msg = "Make sure to name the column to contain the moon information \"moon\".",
                incorrect_msg = "The \"moon\" column does not contain the correct information. Try again.")

test_data_frame(name = "planets_df",
                columns = "mass",
                undefined_cols_msg = "Make sure to name the column to contain the mass information \"mass\".",
                incorrect_msg = "The \"mass\" column does not contain the correct information. Try again.")

test_object("planets_df", incorrect_msg = "It appears that you have correctly specified the \"moon\" and \"mass\" columns, but there's still something wrong with the resulting `planets_df`. Make sure to add columns twice!")
success_msg("Nice one! This data frame is beginning to contain quite some information!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:8e5ade7078
## Sorting

In data analysis you will often sort your data according to a certain variable in the data set. In R, this is done with the help of the function [`order()`](http://www.rdocumentation.org/packages/base/functions/order). 

[`order()`](http://www.rdocumentation.org/packages/base/functions/order) is a function that gives you the ranked position of each element when it is applied on a variable, such as a vector for example:

```
a <- c(100,9,101)
order(a)
``` 

this code returns the vector containing 2, 1 and 3; that's because 100 is the second largest element of the vector, 9 is the smallest element and 101 is the largest element.
    
```
a[order(a)]
```

will thus give you the ordered vector (9, 100, 101), since it first picks the second element of `a`, then the first and then the last. Got it? If you are not sure, use the console to play with the [`order()`](http://www.rdocumentation.org/packages/base/functions/order) function. 

*** =instructions
Experiment with the [`order()`](http://www.rdocumentation.org/packages/base/functions/order) function in the console. Click 'Submit Answer' when you are ready to continue.

*** =hint
Just play with the [`order()`](http://www.rdocumentation.org/packages/base/functions/order) function in the console!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Just play around with the order function in the console to see how it works!
```

*** =solution
```{r}
# Just play around with the order function in the console to see how it works!
# Some examples:
order(1:10)
order(2:11)
order(c(5,4,6,7))
```

*** =sct
```{r}
success_msg("Great! Now let's use the [`order()`](http://www.rdocumentation.org/packages/base/functions/order) function to sort your data frame!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ec87541ef1
## Sorting your data frame

Alright, now let us do something useful with the [`order()`](http://www.rdocumentation.org/packages/base/functions/order) function! You would like to rearrange your data frame such that it starts with the smallest planet and ends with the largest one. A sort on the `diameter` column, if you will.

Suppose you have a data frame `df`, with three columns `a`, `b` and `c`. The following code will print a version of df that is sorted on the column `a`.

```
pos <- order(df$a)
df[pos, ]
```

*** =instructions
- Assign to the variable `positions` the desired ordering for the new data frame that you will create in the next step. You can use the [`order()`](http://www.rdocumentation.org/packages/base/functions/order) function for that.
- Now create the data frame `smallest_first_df`, which contains the same information as `planets_df`, but with the planets in increasing order of magnitude. Use the previously created variable `positions` as row indices inside square brackets to achieve this.
- Print `smallest_first_df` to see what you've accomplished.

*** =hint
- `order(planets_df$diameter)` will give you the ordering of the variable diameter from smallest to largest. This is what you should assign to `positions`.
- Use the variable positions then to select from the data frame `planets_df`: `planets_df[positions, ]`.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter6.RData"))
```

*** =sample_code
```{r}
# planets_df is pre-loaded in your workspace

# Create a desired ordering for planets_df: positions


# Create a new, ordered data frame: smallest_first_df


# Print smallest_first_df
```

*** =solution
```{r}
# planets_df is pre-loaded in your workspace

# Create a desired ordering for planets_df: positions
positions <- order(planets_df$diameter)

# Create a new, ordered data frame: smallest_first_df
smallest_first_df <- planets_df[positions, ]

# Print smallest_first_df
smallest_first_df
```

*** =sct
```{r}
msg = "Do not remove or overwrite the `planets_df` data frame!"
test_object("planets_df", undefined_msg = msg, incorrect_msg = msg)
test_object("positions",
            incorrect_msg = "It looks like `positions` does not contain all the correct ordering of the diameter column.")
test_object("smallest_first_df",
            incorrect_msg = "It looks like `smallest_first_df` does not contain the positions of the ordered `planets_df`.")
test_output_contains("smallest_first_df", incorrect_msg = "Finish off by printing `smallest_first_df`.")
success_msg("Wonderful! What does the resulting data frame look like? Order prevailed!")
```

