--- 
title_meta  : Chapter 1
title       : Intro to basics
description : "In this chapter, you will take your first steps with R. You will learn how to use the console as a calculator and how to assign variables. You will also get to know the basic data types in R. Let us get started!"
attachments : 
  slides_link: https://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/slides/ch1_slides_v2.pdf
  
--- type:VideoExercise xp:50 skills:1
## Meet R

*** =video_link
```{r,eval=FALSE}
//player.vimeo.com/video/144351865
```

*** =video_stream
```{r,eval=FALSE}
https://player.vimeo.com/external/144351865.hd.mp4?s=9956831b02e36428470cd9bdce79bdbc&profile_id=119
```

*** =video_hls
```{r,eval=FALSE}
https://player.vimeo.com/external/144351865.m3u8?p=high,standard,mobile&s=211e71cbc69a0d43b11ab1a1b9d3c7aa
```

--- type:NormalExercise xp:100 skills:1
## Experiment in the console

For this first interactive exercise, simply experiment with code in the console, on the bottom right. 

Try, for example, to type in

```
3 * 4
```

and hit Enter. R should print out

```
[1] 12
```

*** =instructions
If you're done experimenting, click _Submit Answer_.

*** =hint
Simply experiment in the console on the bottom right by typing in simple calculations.

*** =pre_exercise_code
```{r eval=FALSE}
```

*** =sample_code
```{r eval=FALSE}
```

*** =solution
```{r eval=FALSE}
```

*** =sct
```{r eval=FALSE}
success_msg("Great!")
```

--- type:VideoExercise xp:50 skills:1
## Variables and scripts

*** =video_link
```{r,eval=FALSE}
//player.vimeo.com/video/144351866
```

*** =video_stream
```{r,eval=FALSE}
https://player.vimeo.com/external/144351866.hd.mp4?s=692f6fcc48e294ae10ab9f0f30876ab5&profile_id=119
```

*** =video_hls
```{r,eval=FALSE}
https://player.vimeo.com/external/144351866.m3u8?p=high,standard,mobile&s=92850798bd07b64f778d571bd3683363
```

--- type:NormalExercise xp:100 skills:1
## Your first R script

In the script on the right you should type R code to solve the exercises. When you hit the _Submit Answer_ button, every line of code in the script is interpreted and executed by R and you get a message that indicates whether or not your code was correct. The output of your submission is shown in the R console.

You can also execute R commands straight in the console. When you type in the console, your submission will not be checked for correctness!

*** =instructions
In the editor, add another line of code that calculates the sum of 6 and 12, and hit the _Submit Answer_ button.

*** =hint
Simply add a line of R code that calculates the sum of 6 and 12, just like the example in the sample code!

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
3 + 4
```

*** =solution
```{r eval=FALSE}
3 + 4
6 + 12
```

*** =sct
```{r eval=FALSE}
test_output_contains("18", incorrect_msg = "Make sure to add a line of R code, that calculates the sum of 6 and 12.")
success_msg("Awesome! See how the console shows the result of the R code you submitted?")
```


--- type:NormalExercise xp:100 skills:1
## Documenting your code

Adding comments to your code is extremely important to make sure that you and others can understand what your code is about. R makes use of the `#` sign to add comments, just like Twitter!

Comments are not run as R code, so they will not influence your result. For example, _Calculate 3 + 4_ in the editor on the right is a comment and is ignored during execution.

*** =instructions
Add another comment in the editor on the right, _Calculate 6 + 12_, at the appropriate location.

*** =hint
Simply add the line `# Calculate 6 + 12` above the R code that calculates 6 + 12.

*** =pre_exercise_coe
```{r}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Calculate 3 + 4
3 + 4


6 + 12
```

*** =solution_code
```{r}
# Calculate 3 + 4
3 + 4

# Calculate 6 + 12
6 + 12
```

*** =sct
```{r}
test_output_contains("7", incorrect_msg = "Do not remove the code that calculates 3 + 4.")
test_student_typed("# Calculate 3 + 4", not_typed_msg = "Do not remove the comment for the code that calculates 3 + 4.")
test_output_contains("18", incorrect_msg = "Do not remove the code that calculates 6 + 12.")
test_student_typed(c("# Calculate 6 + 12", "# calculate 6 + 12", "#Calculate 6 + 12", "#calculate 6 + 12",
                     "# Calculate 6+12", "# calculate 6+12", "#Calculate 6+12", "#calculate 6+12"), 
                   not_typed_msg = "Make sure to add the comment: `# Calculate 6 + 12`")
success_msg("Great! Looks better, doesn't it?")
```

--- type:NormalExercise xp:100 skills:1
## R as a calculator

In its most basic form R can be used as a scientific calculator. Consider the following arithmetic operators:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`
- Modulo: `%%`

The last two might need some explaining:
- The `^` operator raises the number to its left to the power of the number to its right: for example `3^2` equals 9.
- The modulo returns the remainder of the division of the number to the left by the number on its right, for example 5 modulo 3 or `5 %% 3` equals 2.

*** =instructions
- Type `2^5` in the editor to calculate 2 to the power 5.
- Type `28 %% 6` to calculate 28 modulo 6.
- Click _Submit Answer_ and have a look at the R output in the console.

*** =hint
Another example of the modulo operator: `9 %% 2` equals `1`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Addition
5 + 5 

# Subtraction
5 - 5 

# Multiplication
3 * 5

 # Division
(5 + 5) / 2 

# Exponentiation


# Modulo

```

*** =solution
```{r eval=FALSE}
# Addition
5 + 5 

# Subtraction
5 - 5 

# Multiplication
3 * 5

 # Division
(5 + 5) / 2 

# Exponentiation
2 ^ 5

# Modulo
28 %% 6
```

*** =sct
```{r eval=FALSE}

msg <- "Do not remove the examples that have already been coded for you!"
test_output_contains("5 + 5", incorrect_msg = msg)
test_output_contains("5 - 5", incorrect_msg = msg)
test_output_contains("3 * 5", incorrect_msg = msg)
test_output_contains("(5 + 5)/2", incorrect_msg = msg)
test_output_contains("2^5", incorrect_msg = "Have another look at the exponentiation. Read the instructions carefully.")
test_output_contains("28 %% 6", incorrect_msg = "Have another look at the use of the `%%` operator. Read the instructions carefully.")
success_msg("Nice one!")
```

--- type:NormalExercise xp:100 skills:1
## Variable assignment

A variable allows you store a value or an object in R. You can then later use this variable's name to easily access the value or the object that is stored within this variable. You use `<-` to assign a variable:

```
my_variable <- 4
```

Suppose you have a fruit basket with five apples. You want to store the number of apples in a variable with the name `my_apples`. 

*** =instructions
- Using `<-`, assign the value 5 to `my_apples` below the first comment.
- Type `my_apples` below the second comment. This will print out the value of `my_apples`.
- After clicking _Submit Answer_, have a look at the console: the number 5 is printed, so R now links the variable `my_apples` to the value 5.

*** =hint
Remember that if you want to assign a number or an object to a variable in R, you can make use of the assignment operator `<-`. Alternatively, you can use `=`, but `<-` is widely preferred in the R community.

*** =pre_exercise_code
```{r eval=FALSE}
```

*** =sample_code
```{r eval=FALSE}
# Assign the value 5 to the variable called my_apples


# Print out the value of the variable my_apples

```

*** =solution
```{r eval=FALSE}
# Assign the value 5 to the variable called my_apples
my_apples <- 5

# Print out the value of the variable my_apples
my_apples
```

*** =sct
```{r eval=FALSE}
test_object("my_apples", incorrect_msg = "Have you correctly assigned 5 to `my_apples`? Write `my_apples <- 5` on a new line in the script.")
test_output_contains("my_apples", incorrect_msg = "Have you explicitly told R to print out the `my_apples` variable to the console? Simply type `my_apples` on a new line.")
success_msg("Great! You could also use `=` for variable assignment, but `<-` is typically prefered.")
```

--- type:NormalExercise xp:100 skills:1
## Variable assignment (2)

Every tasty fruit basket needs oranges, so you decide to add six oranges. You decide to create the variable `my_oranges` and assign the value 6 to it. Next, you want to calculate how many pieces of fruit you have in total. Since you have given meaningful names to these values, you can now code this in a clear way:

```
my_apples + my_oranges
```

*** =instructions
- Assign to `my_oranges` the value 6.
- Add the variables `my_apples` and `my_oranges` and have R simply print the result.
- Combine the variables `my_apples` and `my_oranges` into a new variable `my_fruit`, which is the total amount of fruits in your fruit basket.

*** =hint
`my_fruit` is just the sum of `my_apples` and `my_oranges`. You can use the `+` operator to sum the two and `<-` to assign that value to the variable `my_fruit`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Assign 5 to my_apples
my_apples <- 5

# Assign 6 to my_oranges


# Add my_apples and my_oranges: print the result


# Add my_apples and my_oranges: assign to my_fruit

```

*** =solution
```{r eval=FALSE}
# Assign 5 to my_apples
my_apples <- 5

# Assign 6 to my_oranges
my_oranges <- 6

# Add my_apples and my_oranges: print the result
my_apples + my_oranges

# Add my_apples and my_oranges: assign to my_fruit
my_fruit <- my_apples + my_oranges
```

*** =sct
```{r eval=FALSE}
test_object("my_apples", incorrect_msg = "Do not change the assignment of the `my_apples` variable!")
test_object("my_oranges")
test_output_contains("my_apples + my_oranges",
                     incorrect_msg = "The output does not contain the result of adding `my_apples` and `my_oranges` (second instruction). Try again.")
test_object("my_fruit")
success_msg("Nice one! The great advantage of doing calculations with variables is reusability. If you just change `my_apples` to equal 12 instead of 5 and rerun the script, `my_fruit` will automatically update as well.")
```


--- type:NormalExercise xp:100 skills:1
## The workspace

If you assign a value to a variable, this variable is stored in the workspace. It's the place where all user-defined variables live. The command [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) lists the contents of this workspace. 

`rm(<var_name>)` allows you to remove objects from the workspace again. Try the following code in the console:

```
a <- 1
b <- 2
ls()
rm(a)
ls()
```

The first two lines create the variables `a` and `b`. Calling [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) now shows you that `a` and `b` are in the workspace. After removing `a` with `rm(a)`, the same [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) command will show you that only `b` remains in the workspace. You could also remove both `a` and `b` in a one-liner: `rm(a, b)`.

The first line of the sample code is `rm(list = ls())`. This is a very useful command to clear everything from your workspace!

*** =instructions
- List the contents of the workspace to check that the workspace is empty.
- Create a variable, `horses`, equal to 3, and a variable `dogs`, equal to 7.
- Print out the contents of the workspace again to see that indeed, these two variables are available.
- Remove the `dogs` variable from the workspace.
- Finally, print out the objects in your workspace once more to see that only `horses` and `animals` remain.

*** =hint
All you need is a combination of [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) and [`rm()`](http://www.rdocumentation.org/packages/base/functions/rm) at the right time. Give it a try and let the feedback messages guide you.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Clear the entire workspace
rm(list = ls())

# List the contents of your workspace


# Create the variables horses and dogs


# Inspect the contents of the workspace again


# Remove dogs from the workspace


# Inspect the objects in your workspace once more

```

*** =solution
```{r eval=FALSE}
# Clear the entire workspace
rm(list = ls())

# List the contents of your workspace
ls()

# Create the variables horses and dogs
horses <- 3
dogs <- 7

# Inspect the contents of the workspace again
ls()

# Remove dogs from the workspace
rm(dogs)

# Inspect the objects in your workspace once more
ls()
```

*** =sct
```{r eval=FALSE}
test_output_contains("character(0)", 
                     incorrect_msg = "Make sure to list the contents of your workspace when it is still empty! Don't remove `rm(list = ls())`, which helps you start off with a clean slate if you have been experimenting before.")
test_object("horses")
test_student_typed("dogs <- 7", not_typed_msg = "Make sure to assign the value 7 to a new variable `dogs`.")
test_output_contains('c("dogs", "horses")',
                     incorrect_msg = "Make sure to inspect the contents of the workspace after defining the variables `horses` and `dogs`")
test_function("rm", not_called_msg = "Make sure to call the function `rm()` to remove the variable `horses`.")
test_output_contains('"horses"', incorrect_msg = "Make sure to inspect the objects in your workspace after removing the `dogs` variable.")
success_msg("Awesome! You can now build, inspect and manage your workspace, great!")
```

--- type:VideoExercise xp:50 skills:1
## Basic Data Types

*** =video_link
```{r,eval=FALSE}
//player.vimeo.com/video/138173888
```

*** =video_stream
```{r,eval=FALSE}
https://player.vimeo.com/external/138173888.hd.mp4?s=82fc1b03307e9160d8f00d633623de2d&profile_id=119
```

*** =video_hls
```{r,eval=FALSE}
https://player.vimeo.com/external/138173888.m3u8?p=high,standard,mobile&s=2fa96fc053640c9181af7546b03f6451
```

--- type:NormalExercise xp:100 skills:1
## Discover Basic Data Types

Some of R's most basic types to get started are:

- Decimals values like `4.5` are called **numerics**.
- Natural numbers like `4L` are called **integers**. Integers are also numerics.
- Boolean values (`TRUE` or `FALSE`) are called **logical**.
- Text (or string) values are called **characters**.

Note how the quotation marks on the right indicate that `"some text"` is a character.

*** =instructions
Change the value of the:
- `my_numeric` variable to `42`.
- `my_character` variable to `"forty-two"`. Note that the quotation marks indicate that `"forty-two"` is a character.
- `my_logical` variable to `FALSE`.

*** =hint 
Replace the values in the editor with the values that are provided in the exercise.
```
my_numeric <- 42
```
assigns the value 42 to the variable `my_numeric`. 

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# What is the answer to the universe?
my_numeric <- 42.5

# The quotation marks indicate that the variable is of type character
my_character <- "some text"

# Change the value of my_logical
my_logical <- TRUE
```

*** =solution
```{r eval=FALSE}
# What is the answer to the universe?
my_numeric <- 42

# The quotation marks indicate that the variable is of type character
my_character <- "forty-two"

# Change the value of my_logical
my_logical <- FALSE
```

*** =sct
```{r eval=FALSE}
test_object("my_numeric", 
            incorrect_msg = "Make sure that you assign the correct value to `my_numeric.`")
test_object("my_character",
            incorrect_msg = paste("Make sure that you assign the correct value to `my_character`.",
                                  "Do not forget the quotes and beware of capitalization! R is case sensitive!"))
test_object("my_logical",
            undefined_msg = "Please make sure to define a variable `my_logical`.",
            incorrect_msg = "Make sure that you assign the correct value to `my_logical`.") 
success_msg("Great work! Continue to the next exercise.")
```

--- type:NormalExercise xp:100 skills:1
## Back to Apples and Oranges

Common knowledge tells you not to add apples and oranges. But hey, that is what you just did! The `my_apples` and `my_oranges` variables both contained a number in the previous exercise. The `+` operator works with numeric variables in R. 

However, If you try to add a numeric and a character string, R will complain.

*** =instructions
- Click _Submit Answer_ and read the error message. Make sure you understand why this did not work.
- Adjust `my_oranges <- "six"` such that R knows you have 6 oranges and thus a fruit basket with 11 pieces of fruit. Click _Submit Answer_ again.

*** =hint
You have to assign the numeric value `6` to the `my_oranges` variable instead of the character value `"six"`. Notice how the quotation marks are used to indicate that `"six"` is a character.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Assign a value to the variable my_apples and print it out
my_apples <- 5
my_apples       

# Assign a value to the variable my_oranges and print it out
my_oranges <- "six" 
my_oranges 

# New variable that contains the total amount of fruit
my_fruit <- my_apples + my_oranges 
my_fruit
```

*** =solution
```{r eval=FALSE}
# Assign a value to the variable my_apples and print it out
my_apples <- 5  
my_apples  

# Assign a value to the variable my_oranges and print it out
my_oranges <- 6
my_oranges 

# New variable that contains the total amount of fruit
my_fruit <- my_apples + my_oranges 
my_fruit
```

*** =sct
```{r eval=FALSE}
test_object("my_apples", incorrect_msg = "Don't change the code that assigns 5 to `my_apples`.")
test_object("my_oranges", incorrect_msg = "Change the assignment of the `my_oranges` variable such that the code runs without errors.")
test_object("my_fruit", 
            undefined_msg = "Please make sure to define a variable `my_fruit`.",
            incorrect_msg = "Make sure that you assign the correct value to `my_fruit`.")
test_output_contains("my_fruit", incorrect_msg = "The output does not contain the result of adding `my_apples` and `my_oranges`.")
success_msg("Awesome, keep up the good work!")
```

--- type:MultipleChoiceExercise xp:50
## What's that data type?

When you added the variables containing `5` and `"six"`, you got an error due to a mismatch in data types. You can avoid such embarrassing situations by checking the data type of a variable beforehand:

```
class(<my_var>)
```

In the workspace (you can see what it contains by typing [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) in the console), some variables have already been defined. Which statement concerning these variables are correct?

*** =instructions
- `a`'s class is `integer`, `b` is a `character`, `c` is a `boolean`.
- `a`'s class is `character`, `b` is a `character` as well, `c` is a `logical`.
- `a`'s class is `numeric`, `b` is a `string`, `c` is a `logical`.
- `a`'s class is `numeric`, `b` is a `character`, `c` is a `logical`.

*** =hint
You can find out the data type of the `a` variable for example by typing `class(a)`. You can do similar things for `b` and `c`.

*** =pre_exercise_code
```{r eval=FALSE}
a <- 42
b <- "forty-two"
c <- FALSE
```

*** =sct
```{r eval=FALSE}
msg1 <- "`boolean` is not the class for logical values. Try again."
msg2 <- "`a` is of the class `numeric`, give it another go."
msg3 <- "`string` is not a class in R. `character` is!"
msg4 <- "Nice one. Let's step it up a notch and start coercing variables!"
test_mc(correct = 4, feedback_msgs = c(msg1, msg2, msg3, msg4))
```

--- type:NormalExercise xp:100 skills:1
## Coercion: Taming your data

As Filip explained in the video, coercion to transform your data from one type to another is perfectly possible. Next to the [`class()`](http://www.rdocumentation.org/packages/base/functions/class) function and the `is.*()` functions, you can use the `as.*()` functions to enforce data to change types. For example, 

```
var <- "3"
var_num <- as.numeric(var)
```

converts the character string `"3"` in `var` to a numeric `3` and assigns it to `var_num`. Beware however, that it is not always possible to convert the types without information loss or errors:

```
as.integer("4.5")
as.numeric("three")
```

The first line will convert the character string `"4.5"` to the integer `4`. The second one will convert the character string `"three"` to an `NA`.

*** =instructions
- Convert `var1`, a logical, to a character and assign it to the variable `var1_char`.
- Inspect the closs of `var1_char` by using [`class()`](http://www.rdocumentation.org/packages/base/functions/class) on it.
- Try to coerce `var2` to a numeric and assign the result to `var2_num`. Was it successful?

*** =hints
Use the [`as.character()`](http://www.rdocumentation.org/packages/base/functions/character) function to convert `var1` to a character. For `var2`, use [`as.numeric()`](http://www.rdocumentation.org/packages/base/functions/numeric).

*** =sample_code
```{r}
# Create variables var1, var2
var1 <- TRUE
var2 <- "i"

# Convert var1 to a character: var1_char


# Display the class of var1_char


# Coerce var2 to a numeric: var2_num
  
```

*** =solution
```{r}
# Create variables var1, var2
var1 <- TRUE
var2 <- "i"

# Convert var1 to a character: var1_char
var1_char <- as.character(var1)

# Display the class of var1_char
class(var1_char)

# Coerce var2 to a numeric: var2_num
var2_num <- as.numeric(var2)
```

*** =sct
```{r}
msg <- "Do not remove or change the definition of the variables `var1`, `var2` and `var3`."
test_object("var1", undefined_msg = msg, incorrect_msg = msg)
test_object("var2", undefined_msg = msg, incorrect_msg = msg)
test_function("as.character", "x",
              not_called_msg = "Make sure to call the function `as.character()` to convert `var1` to a character.",
              incorrect_msg = "Have you passed the correct variable to the function `as.character()`?")
test_object("var1_char")
test_function("class", "x", 
              not_called_msg = "Make sure to call the function `class()` to inspect the class of `var1_char()`.",
              incorrect_msg = "Have you passed the correct variable to the function `class()`?")
test_function("as.numeric", "x",
              not_called_msg = "Make sure to call the function `as.numeric()` to convert `var2` to a numeric.",
              incorrect_msg = "Have you passed the correct variable to the function `as.numeric()`?")
test_object("var2_num")
success_msg("Bellissimo! The coercion of `var2` did not succeed, hence the warning.")
```
