--- 
title_meta  : Chapter 1
title       : Intro to basics
description : "In this chapter, you will take your first steps with R. You will learn how to use the console as a calculator and how to assign variables. You will also get to know the basic data types in R. Let's get started!"
attachments : 
  slides_link: https://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/slides/ch1_slides_v2.pdf
  
--- type:VideoExercise lang:r xp:50 skills:1 key:1a1ba28cd5
## Meet R

*** =video_link
//player.vimeo.com/video/144351865

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch1_1.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:5714863ba8
## Your first R script

In the script on the right you should type R code to solve the exercises. When you hit the _Submit Answer_ button, every line of code in the script is interpreted and executed by R and you get a message that indicates whether or not your code was correct. The output of your submission is shown in the R console.

You can also execute R commands straight in the console. When you type in the console, your submission will not be checked for correctness! Try, for example, to type in `3 * 4` and hit Enter. R should return `[1] 12`.

*** =instructions
In the script, add another line of code that calculates the sum of 6 and 12, and hit the _Submit Answer_ button.

*** =hint
Simply add a line of R code that calculates the sum of 6 and 12, just like the example in the sample code!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
3 + 4
```

*** =solution
```{r}
3 + 4
6 + 12
```

*** =sct
```{r}
test_output_contains("18", incorrect_msg = "Make sure to add a line of R code that calculates the sum of 6 and 12.")
success_msg("Awesome! See how the console shows the result of the R code you submitted?")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ab37530088
## Documenting your code

Adding comments to your code is extremely important to make sure that you and others can understand what your code is about. R makes use of the `#` sign to add comments, just like Twitter!

Comments are not run as R code, so they will not influence your result. For example, _Calculate 3 + 4_ in the script on the right is a comment and is ignored during execution.

*** =instructions
Add another comment in the script on the right, _Calculate 6 + 12_, at the appropriate location.

*** =hint
Simply add the line `# Calculate 6 + 12` above the R code that calculates 6 + 12.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Calculate 3 + 4
3 + 4


6 + 12
```

*** =solution
```{r}
# Calculate 3 + 4
3 + 4

# Calculate 6 + 12
6 + 12
```

*** =sct
```{r}
test_output_contains("7", incorrect_msg = "Do not remove the code that calculates 3 + 4.")
test_student_typed("# Calculate 3 + 4", not_typed_msg = "Do not remove the comment for the code that calculates 3 + 4.")
test_output_contains("18", incorrect_msg = "Do not remove the code that calculates 6 + 12.")
test_student_typed(c("# Calculate 6 + 12", "# calculate 6 + 12", "#Calculate 6 + 12", "#calculate 6 + 12",
                     "# Calculate 6+12", "# calculate 6+12", "#Calculate 6+12", "#calculate 6+12"), 
                   not_typed_msg = "Make sure to add the comment: `# Calculate 6 + 12`")
success_msg("Great! Looks better, doesn't it?")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:9d8a3d0b88
## R as a calculator

In its most basic form R can be used as a scientific calculator. Consider the following arithmetic operators:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`
- Modulo: `%%`

The last two might need some explaining:
- The `^` operator raises the number to its left to the power of the number to its right: for example `3^2` equals 9.
- The modulo returns the remainder of the division of the number to the left by the number on its right, for example 5 modulo 3 or `5 %% 3` equals 2.

*** =instructions
- Type `2^5` in the script to calculate 2 to the power 5.
- Type `28 %% 6` to calculate 28 modulo 6.
- Click _Submit Answer_ and have a look at the R output in the console.

*** =hint
Another example of the modulo operator: `9 %% 2` equals `1`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Addition
5 + 5 

# Subtraction
5 - 5 

# Multiplication
3 * 5

 # Division
(5 + 5) / 2 

# Exponentiation


# Modulo

```

*** =solution
```{r}
# Addition
5 + 5 

# Subtraction
5 - 5 

# Multiplication
3 * 5

 # Division
(5 + 5) / 2 

# Exponentiation
2 ^ 5

# Modulo
28 %% 6
```

*** =sct
```{r}
msg <- "Do not remove the examples that have already been coded for you!"
test_output_contains("5 + 5", incorrect_msg = msg)
test_output_contains("5 - 5", incorrect_msg = msg)
test_output_contains("3 * 5", incorrect_msg = msg)
test_output_contains("(5 + 5)/2", incorrect_msg = msg)
test_output_contains("2^5", incorrect_msg = "Have another look at the exponentiation. Read the instructions carefully.")
test_output_contains("28 %% 6", incorrect_msg = "Have another look at the use of the `%%` operator. Read the instructions carefully.")
success_msg("Nice one!")
```


--- type:MultipleChoiceExercise xp:50 skills:1 key:9d8819fb2e
## R's pros and cons

As Filip explained in the video, there are things that make R the awesome and immensely popular language that it is today. On the other hand, there are also aspects about R that are less attractive. Which of the following statements are true regarding this statistical programming language developed by Ihaka and Gentleman in the nineties?

1. As opposed to SAS and SPSS, R is completely open-source.
2. R is open-source, but it's hard to share your code with others since R uses a command-line interface.
3. It typically takes a long time for new and updated R packages to be released and made available to the public.
4. R is easy to use, but this comes at the cost of limited graphical abilities.
5. R works well with large data sets, if the code is properly written and the data fits into the working memory. 

*** =instructions
- statements (1) and (2) are correct; the others are false.
- statements (1) and (4) are correct; the others are false.
- statements (1) and (5) are correct; the others are false.
- statements (2) and (4) are correct; the others are false.
- statements (3) and (5) are correct; the others are false.

*** =hint
Remember that your data has to fit in the working memory for R to be able to process it.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sct
```{r}
msg1 = "Remember that the fact that R uses a command-line interface, does not make it hard to share code. On the contrary, sharing your results becomes very straightforward because you can easily share R scripts."
msg2 = "R is the perfect tool for creating neat and insightful visualizations. Try again."
msg3 = "Great! Head over to the next exercise and get your hands dirty!"
msg4 = "R uses a command-line interface, which makes it very easy to share one's code. Also, R is very suitable for creating visualizations. Try again."
msg5 = "It's fairly straightforward to write, maintain and share R packages. Try again."
test_mc(3, feedback_msgs = c(msg1, msg2, msg3, msg4, msg5))
```


--- type:NormalExercise lang:r xp:100 skills:1 key:6b6fb4974c
## Variable assignment (1)

A variable allows you store a value or an object in R. You can then later use this variable's name to easily access the value or the object that is stored within this variable. You use `<-` to assign a variable:

```
my_variable <- 4
```

*** =instructions
Complete the code in the editor such that it assigns the value 42 to the variable `x` in the editor. Click 'Submit Answer'. Notice that when you ask R to print `x`, the value 42 appears.

*** =hint
Look at how the value 4 was assigned to `my_variable` in the exercise's assignment. Do the exact same thing in the editor, but now assign 42 to the variable `x`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Assign the value 42 to x
x <- 

# Print out the value of the variable x
x
```

*** =solution
```{r}
# Assign the value 42 to x
x <- 42

# Print out the value of the variable x
x
```

*** =sct
```{r}
test_error()
test_object("x", 
            undefined_msg = "Make sure to define a variable <code>x</code>.",
            incorrect_msg = "Make sure that you assign the correct value to <code>x</code>.") 
success_msg("Good job! Notice that R does not print the value of a variable to the console when you do the assignment. <code>x <- 42</code> did not generate any output, because R assumes that you will be needing this variable in the future. Otherwise you wouldn't have stored the value in a variable in the first place, right? Proceed to the next exercise!")
```




--- type:NormalExercise xp:100 skills:1 key:a5b8028834
## Variable assignment (2)

Suppose you have a fruit basket with five apples. You want to store the number of apples in a variable with the name `my_apples`. 

*** =instructions
- Using `<-`, assign the value 5 to `my_apples` below the first comment.
- Type `my_apples` below the second comment. This will print out the value of `my_apples`.
- After clicking _Submit Answer_, have a look at the console: the number 5 is printed, so R now links the variable `my_apples` to the value 5.

*** =hint
Remember that if you want to assign a number or an object to a variable in R, you can make use of the assignment operator `<-`. Alternatively, you can use `=`, but `<-` is widely preferred in the R community.

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# Assign the value 5 to the variable called my_apples


# Print out the value of the variable my_apples

```

*** =solution
```{r}
# Assign the value 5 to the variable called my_apples
my_apples <- 5

# Print out the value of the variable my_apples
my_apples
```

*** =sct
```{r}
test_object("my_apples", incorrect_msg = "Have you correctly assigned 5 to `my_apples`? Write `my_apples <- 5` on a new line in the script.")
test_output_contains("my_apples", incorrect_msg = "Have you explicitly told R to print out the `my_apples` variable to the console? Simply type `my_apples` on a new line.")
success_msg("Great! You could also use `=` for variable assignment, but `<-` is typically preferred.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:a0cb1bea96
## Variable assignment (3)

Every tasty fruit basket needs oranges, so you decide to add six oranges. You decide to create the variable `my_oranges` and assign the value 6 to it. Next, you want to calculate how many pieces of fruit you have in total. Since you have given meaningful names to these values, you can now code this in a clear way:

```
my_apples + my_oranges
```

*** =instructions
- Assign to `my_oranges` the value 6.
- Add the variables `my_apples` and `my_oranges` and have R simply print the result.
- Combine the variables `my_apples` and `my_oranges` into a new variable `my_fruit`, which is the total amount of fruits in your fruit basket.

*** =hint
`my_fruit` is just the sum of `my_apples` and `my_oranges`. You can use the `+` operator to sum the two and `<-` to assign that value to the variable `my_fruit`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Assign 5 to my_apples
my_apples <- 5

# Assign 6 to my_oranges


# Add my_apples and my_oranges: print the result


# Add my_apples and my_oranges: assign to my_fruit

```

*** =solution
```{r}
# Assign 5 to my_apples
my_apples <- 5

# Assign 6 to my_oranges
my_oranges <- 6

# Add my_apples and my_oranges: print the result
my_apples + my_oranges

# Add my_apples and my_oranges: assign to my_fruit
my_fruit <- my_apples + my_oranges
```

*** =sct
```{r}
test_object("my_apples", incorrect_msg = "Do not change the assignment of the `my_apples` variable!")
test_object("my_oranges")
test_output_contains("my_apples + my_oranges",
                     incorrect_msg = "The output does not contain the result of adding `my_apples` and `my_oranges` (second instruction). Try again.")
test_object("my_fruit")
success_msg("Nice one! The great advantage of doing calculations with variables is reusability. If you just change `my_apples` to equal 12 instead of 5 and rerun the script, `my_fruit` will automatically update as well.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:6192f64167
## The workspace

If you assign a value to a variable, this variable is stored in the workspace. It's the place where all user-defined variables live. The command [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) lists the contents of this workspace. 

```
a <- 1
b <- 2
ls()
```

The first two lines create the variables `a` and `b`. Calling [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) now shows you that `a` and `b` are in the workspace. 

You can also remove variables from the workspace. You do this with [`rm()`](http://www.rdocumentation.org/packages/base/functions/rm). `rm(a)`, for example, would remove `a` from the workspace again. `rm(list = ls())`, which is used in the beginning of your script, clears everything from the workspace.

*** =instructions
- Create a variable, `horses`, equal to 3, and a variable `dogs`, equal to 7.
- List the contents of your workspace with [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) to see that indeed, these two variables are in there.

*** =hint
All you need is a combination of [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) and [`rm()`](http://www.rdocumentation.org/packages/base/functions/rm) at the right time. Give it a try and let the feedback messages guide you.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Clear the entire workspace
rm(list = ls())

# Create the variables horses and dogs


# List the contents of your workspace


```

*** =solution
```{r}
# Clear the entire workspace
rm(list = ls())

# Create the variables horses and dogs
horses <- 3
dogs <- 7

# Inspect the contents of the workspace again
ls()
```

*** =sct
```{r}
test_student_typed("rm(list = ls())", not_typed_msg = "Do not remove the line `rm(list = ls())`.")
test_object("horses")
test_object("dogs")
test_output_contains('c("dogs", "horses")',
                     incorrect_msg = "Make sure to inspect the objects in your workspace after creating `horses` and `dogs`.")
success_msg("Awesome! You can now build up and inspect your workspace, great!")
```


--- type:VideoExercise lang:r xp:50 skills:1 key:9f9019501e
## Basic Data Types

*** =video_link
//player.vimeo.com/video/138173888

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch1_2.master.m3u8



--- type:NormalExercise lang:r xp:100 skills:1 key:1866cdd202
## Discover Basic Data Types

To get started, here are some of R's most basic data types:

- Decimal values like `4.5` are called **numerics**.
- Natural numbers like `4L` are called **integers**. Integers are also numerics.
- Boolean values (`TRUE` or `FALSE`) are called **logical**. Capital letters are important here; `true` and `false` are not valid.
- Text (or string) values are called **characters**.

Note how the quotation marks on the right indicate that `"some text"` is of type character.

*** =instructions
Change the value of the:

- `my_numeric` variable to `42`.
- `my_character` variable to `"forty-two"`. Note that the quotation marks indicate that `"forty-two"` is a character.
- `my_logical` variable to `FALSE`.

*** =hint 
Replace the values in the script with the values that are provided in the exercise.
```
my_numeric <- 42
```
assigns the value 42 to the variable `my_numeric`. 

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# What is the answer to the universe?
my_numeric <- 42.5

# The quotation marks indicate that the variable is of type character
my_character <- "some text"

# Change the value of my_logical
my_logical <- TRUE
```

*** =solution
```{r}
# What is the answer to the universe?
my_numeric <- 42

# The quotation marks indicate that the variable is of type character
my_character <- "forty-two"

# Change the value of my_logical
my_logical <- FALSE
```

*** =sct
```{r}
test_object("my_numeric", 
            incorrect_msg = "Make sure that you assign the correct value to `my_numeric.`")
test_object("my_character",
            incorrect_msg = paste("Make sure that you assign the correct value to `my_character`.",
                                  "Do not forget the quotes and beware of capitalization! R is case sensitive!"))
test_object("my_logical",
            undefined_msg = "Please make sure to define a variable `my_logical`.",
            incorrect_msg = "Make sure that you assign the correct value to `my_logical`.") 
success_msg("Great work! Continue to the next exercise.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:c52153af0b
## Back to Apples and Oranges

Common knowledge tells you not to add apples and oranges. But hey, that is what you just did! The `my_apples` and `my_oranges` variables both contained a number in the previous exercise. The `+` operator works with numeric variables in R. 

However, if you try to add a numeric and a character string, R will complain.

*** =instructions
- Click _Submit Answer_ and read the error message. Make sure you understand why this did not work.
- Adjust `my_oranges <- "six"` such that R knows you have 6 oranges and thus a fruit basket with 11 pieces of fruit. Click _Submit Answer_ again.

*** =hint
You have to assign the numeric value `6` to the `my_oranges` variable instead of the character value `"six"`. Notice how the quotation marks are used to indicate that `"six"` is a character.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Assign a value to the variable my_apples and print it out
my_apples <- 5
my_apples       

# Assign a value to the variable my_oranges and print it out
my_oranges <- "six" 
my_oranges 

# New variable that contains the total amount of fruit
my_fruit <- my_apples + my_oranges 
my_fruit
```

*** =solution
```{r}
# Assign a value to the variable my_apples and print it out
my_apples <- 5  
my_apples  

# Assign a value to the variable my_oranges and print it out
my_oranges <- 6
my_oranges 

# New variable that contains the total amount of fruit
my_fruit <- my_apples + my_oranges 
my_fruit
```

*** =sct
```{r}
test_object("my_apples", incorrect_msg = "Don't change the code that assigns 5 to `my_apples`.")
test_object("my_oranges", incorrect_msg = "Change the assignment of the `my_oranges` variable such that the code runs without errors.")
test_object("my_fruit", 
            undefined_msg = "Please make sure to define a variable `my_fruit`.",
            incorrect_msg = "Make sure that you assign the correct value to `my_fruit`.")
test_output_contains("my_fruit", incorrect_msg = "The output does not contain the result of adding `my_apples` and `my_oranges`.")
success_msg("Awesome, keep up the good work!")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:7806ca24d2
## What's that data type?

When you added the variables containing `5` and `"six"`, you got an error due to a mismatch in data types. You can avoid such embarrassing situations by checking the data type of a variable beforehand:

```
class(my_var)
```

In the workspace (you can see what it contains by typing [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls) in the console), some variables have already been defined. Which statement concerning these variables are correct?

*** =instructions
- `a`'s class is `integer`, `b` is a `character`, `c` is a `boolean`.
- `a`'s class is `character`, `b` is a `character` as well, `c` is a `logical`.
- `a`'s class is `numeric`, `b` is a `string`, `c` is a `logical`.
- `a`'s class is `numeric`, `b` is a `character`, `c` is a `logical`.

*** =hint
You can find out the data type of the `a` variable for example by typing `class(a)`. You can do similar things for `b` and `c`.

*** =pre_exercise_code
```{r}
a <- 42
b <- "forty-two"
c <- FALSE
```

*** =sct
```{r}
msg1 <- "`boolean` is not the class for logical values. Try again."
msg2 <- "`a` is of the class `numeric`, give it another go."
msg3 <- "`string` is not a class in R. `character` is!"
msg4 <- "Nice one. Let's step it up a notch and start coercing variables!"
test_mc(correct = 4, feedback_msgs = c(msg1, msg2, msg3, msg4))
```

--- type:NormalExercise lang:r xp:100 skills:1 key:c75fe45544
## Coercion: Taming your data

As Filip explained in the video, coercion to transform your data from one type to another is perfectly possible. Next to the [`class()`](http://www.rdocumentation.org/packages/base/functions/class) function and the `is.*()` functions, you can use the `as.*()` functions to force data to change types.

Take this example:

```
var <- "3"
var_num <- as.numeric(var)
```

`var`, a character string, is converted into a numeric using [`as.numeric()`](http://www.rdocumentation.org/packages/base/functions/numeric). The resulting numeric is stored as `var_num`.

*** =instructions
- Convert `var`, a logical, to a character. Assign to resulting character string to the variable `var_char`.
- Inspect the class of `var_char` by using [`class()`](http://www.rdocumentation.org/packages/base/functions/class) on it.

*** =hints
Use the [`as.character()`](http://www.rdocumentation.org/packages/base/functions/character) function to convert `var` to a character.

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# Definition of var
var <- TRUE

# Convert var to a character: var_char


# Display the class of var_char


```

*** =solution
```{r}
# Definition of var
var <- TRUE

# Convert var to a character: var_char
var_char <- as.character(var)

# Display the class of var_char
class(var_char)
```

*** =sct
```{r}
test_error()
msg <- "Do not remove or change the definition of the variable `var`."
test_object("var", undefined_msg = msg, incorrect_msg = msg)
test_function("as.character", "x",
              not_called_msg = "Make sure to call the function [`as.character()`](http://www.rdocumentation.org/packages/base/functions/character) to convert `var` to a character.",
              incorrect_msg = "Have you passed the correct variable to the function [`as.character()`](http://www.rdocumentation.org/packages/base/functions/character)?")
test_object("var_char")
test_function("class", "x", 
              not_called_msg = "Make sure to call the function <code>class()</code> to inspect the class of <code>var_char</code>.",
              incorrect_msg = "Have you passed the correct variable to the function <code>class()/<code>?")
success_msg("Bellissimo!")
```
