--- 
title_meta  : Chapter 2
title       : Vectors
description : In this chapter, you will take your first steps with R. You will learn how to use the console as a calculator and how to assign variables. You will also get to know the basic data types in R. Let us get started!
attachments : 
  slides_link: https://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/slides/ch2_slides.pdf
  
--- type:VideoExercise xp:50
## Create and Name Vectors

*** =video_link
```{r,eval=FALSE}
//player.vimeo.com/video/138173896
```

*** =video_stream
```{r,eval=FALSE}
https://player.vimeo.com/external/138173896.hd.mp4?s=ad1bb666a7bd15a35883ae2f912ba5f6&profile_id=119
```

*** =video_hls
```{r,eval=FALSE}
https://player.vimeo.com/external/138173896.m3u8?p=high,standard,mobile&s=63b5262a0f30b79551e2bcb3e484e3b0
```

*** =skills
1

--- type:NormalExercise xp:100
## Create a vector (1)

Feeling lucky? 

You better, because this chapter takes you on a trip to Sin City, also known as *"Statisticians Paradise"* ;-).

Thanks to R and your new data science skills, you will learn how to uplift your performance at the tables and fire off your career as a professional gambler. This chapter will show how you can easily keep track of your betting progress and how you can do some simple analyses on past actions. Next Stop, Vegas Baby... VEGAS!!

On your way from rags to riches, you will make extensive use of vectors. As Filip explained you, vectors are one dimensional arrays that can hold numeric data, character data, or logical data. You create a vector with the combine function [`c()`](http://www.rdocumentation.org/packages/base/functions/c). You place the vector elements separated by a comma between the brackets. For example:

```
numeric_vector <- c(1, 2, 3)
character_vector <- c("a", "b", "c")
boolean_vector <- c(TRUE, FALSE)
```

*** =instructions 
Create a vector, `boolean_vector`, that contains the three elements: `TRUE`, `FALSE` and `TRUE` (in that order). 

*** =hint 
Assign `c(TRUE,FALSE,TRUE)` to the variable `boolean_vector` with the `<-` operator.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")

# Create boolean_vector

```

*** =solution
```{r eval=FALSE}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")

# Create boolean_vector
boolean_vector <- c(TRUE, FALSE, TRUE)
```

*** =sct
```{r eval=FALSE}
test_error()
test_object("numeric_vector")
test_object("character_vector")
test_object("boolean_vector",
            undefined_msg = "Please make sure to define a variable <code>boolean_vector</code>.",
            incorrect_msg = "Make sure that you assign the correct values to <code>boolean_vector</code>. The order matters!")
success_msg("Perfect! Notice that adding a space behind the commas in the <code>c()</code> function improve the readability of your code. Let's practice some more with vector creation in the next exercise.")
```

*** =skills
1

--- type:NormalExercise xp:100
## Create a vector (2)

After one week in Las Vegas and still zero Ferraris in your garage, you decide that it is time to start using your data science superpowers.

Before doing a first analysis, you decide to first collect all the winnings and losses for the last week: 

For `poker_vector`: 
- On Monday you won 140\$ 
- Tuesday you lost 50\$
- Wednesday you won 20\$ 
- Thursday you lost 120\$ 
- Friday you won 240\$

For `roulette_vector`: 
- On Monday you lost 24\$ 
- Tuesday you lost 50\$
- Wednesday you won 100\$ 
- Thursday you lost 350\$
- Friday you won 10\$

You only played poker and roulette, since there was a delegation of mediums that occupied the craps tables. To be able to use this data in R, you decide to create the variables `poker_vector` and `roulette_vector`.

*** =instructions
Assign the winnings/losses for roulette to the variable `roulette_vector`.

*** =hint
To help you with this step, the editor already contains the code for creating `poker_vector`. Assign the correct values to `roulette_vector` based on the numbers in the assignment. Do not forget that losses are negative numbers.

*** =sample_code

```{r eval=FALSE}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday: roulette_vector

```

*** =solution

```{r eval=FALSE}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday: roulette_vector
roulette_vector <- c(-24, -50, 100, -350, 10)
```

*** =sct
```{r eval=FALSE}
test_error()
test_object("poker_vector",
            undefined_msg = "Please make sure to define a variable <code>poker_vector</code>.",
            incorrect_msg = "Make sure that you assign the correct values to <code>poker_vector</code>.")
test_object("roulette_vector",
            undefined_msg = "Please make sure to define a variable <code>roulette_vector</code>.",
            incorrect_msg = "Make sure that you assign the correct values to <code>roulette_vector</code>.")
success_msg("Very good! To check out the contents of your vectors, remember that you can always simply type the variable in the console and hit Enter. Proceed to the next exercise!")
```

*** =skills
1

--- type:NormalExercise xp:100
## Naming a vector (1)

As a data analyst, it is important to have a clear view on the data that you are using. Understanding what each element refers to is therefore essential. 

In the previous exercise, we created a vector with your winnings over the week. Each vector element refers to a day of the week but it is hard to tell which element belongs to which day. It would be nice if you could show that in the vector itself. Remember the [`names()`](http://www.rdocumentation.org/packages/base/functions/names) function to name the elements of a vector?

```
some_vector <- c("Johnny", "Poker Player")
names(some_vector) <- c("Name", "Profession")
```

You can accomplish the exact same thing by using the equals sign inside [`c()`](http://www.rdocumentation.org/packages/base/functions/c):

```
some_vector <- c(Name = "Johnny", Profession = "Poker Player")
```

*** =instructions
Go ahead and assign the days of the week as names to `poker_vector` and `roulette_vector`. In case you are not sure, the days of the week are: Monday, Tuesday, Wednesday, Thursday and Friday.

*** =hint
One way of naming `poker_vector` is by using `names(poker_vector)` and setting this to a vector with the days of the week: 
```
c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```

*** =pre_exercise_code
```{r eval=FALSE}
```

*** =sample_code
```{r eval=FALSE}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Add names to both poker_vector and roulette_vector

```

*** =solution
```{r eval=FALSE}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Add names to both poker_vector and roulette_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(roulette_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```

*** =sct
```{r eval=FALSE}
test_error()
test_object("poker_vector",
            undefined_msg = "Please make sure to define a variable <code>poker_vector</code>.",
            incorrect_msg = "Do not change the values inside <code>poker_vector</code>; they were already coded for you.")
test_object("roulette_vector",
            undefined_msg = "Please make sure to define a variable <code>roulette_vector</code>.",
            incorrect_msg = "Do not change the values inside <code>roulette_vector</code>; they were already coded for you.")
test_object("poker_vector",
            eq_condition = "equal",
            incorrect_msg = "Make sure that you assign the correct names vector to <code>poker_vector</code>. Make sure that the names of the day start with a capital letter!")
test_object("roulette_vector",
            eq_condition = "equal",
            incorrect_msg = "Make sure that you assign the correct names vector to <code>roulette_vector</code>. Make sure that the names of the day start with a capital letter!")
success_msg("Well done! Continue to the next exercise.")
```

*** =skills
1

--- type:NormalExercise xp:100
## Naming a vector (2)

If you want to become a good statistician, you have to become lazy. (If you are already lazy, chances are high you are one of those exceptional, natural-born statistical talents.)

In the previous exercises you probably experienced that it is boring and frustrating to type and retype information such as the days of the week. However, there is a more efficient way to do this, namely, to assign the days of the week vector to a variable! 

Just like you did with your poker and roulette returns, you can also create a variable that contains the days of the week. This way you can use and re-use it.

*** =instructions
- Create a variable `days_vector` that contains the days of the week, from Monday to Friday.
- Use that variable `days_vector` to set the names of `poker_vector` and `roulette_vector`.

*** =hint
You can use `names(poker_vector)` to set the names of the variable `poker_vector`. Make sure not to forget the parentheses around the days of the week.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Create the variable days_vector

 
# Assign the names of the day to roulette_vector and poker_vector

```

*** =solution
```{r eval=FALSE}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Create the variable days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Assign the names of the day to roulette_vector and poker_vector
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
```

*** =sct
```{r eval=FALSE}
test_error()
test_object("poker_vector",
            undefined_msg = "Please make sure to define a variable <code>poker_vector</code>.",
            incorrect_msg = "Do not change the values inside <code>poker_vector</code>; they were already coded for you.")
test_object("roulette_vector",
            undefined_msg = "Please make sure to define a variable <code>roulette_vector</code>.",
            incorrect_msg = "Do not change the values inside <code>roulette_vector</code>; they were already coded for you.")
test_object("days_vector",
            undefined_msg = "Please make sure to define a variable <code>days_vector</code>.",
            incorrect_msg = "Make sure that you assign the correct vector to <code>days_vector</code>.")
test_object("poker_vector",
            eq_condition = "equal",
            incorrect_msg = "Make sure that you assign <code>days_vector</code> to the names of <code>poker_vector</code>.")
test_object("roulette_vector",
            eq_condition = "equal",
            incorrect_msg = "Make sure that you assign <code>days_vector</code> to the names of <code>roulette_vector</code>.")
success_msg("Nice one! A word of advice: try to avoid code duplication at all times. Continue to the next exercise and learn how to do arithmetic with vectors!")
```

*** =skills
1

--- type:MultipleChoiceExercise xp:100
## Different ways to create and name vectors

The previous exercises outlined different ways of creating and naming vectors. Have a look at this chunk of code:

```
poker_vector1 <- c(140, -50, 20, -120, 240)
names(poker_vector1) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

poker_vector2 <- c(Monday = 140, -50, 20, -120, 240)

roulette_vector1 <- c(-24, -50, 100, -350, 10)
days_vector <- names(poker_vector1)
names(roulette_vector1) <- days_vector

roulette_vector2 <- c(-24, -50, 100, -350, 10)
names(roulette_vector2) <- "Monday"
```

Which of the following statements is true?

*** =instructions
- The code to define `poker_vector2` is syntactically invalid.
- `poker_vector1` and `poker_vector2` have different lengths.
- `poker_vector1` and `roulette_vector1` have the same names, while `poker_vector2` and `roulette_vector2` show a names mismatch.
- You can only use [`names()`](http://www.rdocumentation.org/packages/base/functions/names) to _set_ the names of a vector, making `days_vector <- names(poker_vector1)` invalid.

*** =hint
Just try the different code chunks one after the other in the console, and check which statement holds.

*** =pre_exercise_code
```{r}
# NO PEC
```

*** =sct
```{r eval=FALSE}
msg1 <- "The code to define `poker_vector2` is not invalid; try it out in the console!"
msg2 <- "If you use the <code>length()</code> function to inspect the length of both <code>poker_vector1</code> and <code>poker_vector2</code>, you'll see that they both have length 5."
msg3 <- "Correct! You might expect that the names of the vectors <code>roulette_vector1</code> and <code>roulette_vector2</code> are named the same; but the different approaches treat missing name information differently. Also, notice here how you can also use <code>names()</code> to <i>get</i> the names of a vector!"
msg4 <- "Incorrect. You can use <code>names()</code> both to <i>set</i> as to <i>get</i> the names of a vector."
test_mc(3, feedback_msgs = c(msg1, msg2, msg3, msg4))
```

*** =skills
1

--- type:VideoExercise xp:50
## Vector Arithmetic

*** =video_link
```{r,eval=FALSE}
//player.vimeo.com/video/141163398
```

*** =video_stream
```{r,eval=FALSE}
https://player.vimeo.com/external/141163398.hd.mp4?s=89a4fcf84cb5a297243c77db9de7dafb&profile_id=119
```

*** =video_hls
```{r,eval=FALSE}
https://player.vimeo.com/external/141163398.m3u8?p=high,standard,mobile&s=e62fc2293d8e4b3cbe3201cdab3e0639
```

*** =skills
1

--- type:NormalExercise xp:100
## Summing and subtracting vectors

Now that you have the poker and roulette winnings nicely as a named vector, you can start doing some data science magic. 

You want to find out the following type of information:
- How much has been your overall profit or loss per day of the week?
- Have you lost money over the week in total?
- Are you winning/losing money on poker or on roulette?

You'll have to do arithmetic calculations on vectors to solve these problems. Remember that this happens element-wise; the following three statements are completely equivalent:

```
c(1, 2, 3) + c(4, 5, 6)
c(1 + 4, 2 + 5, 3 + 6)
c(5, 7, 9)
```

*** =instructions
- Take the element-wise sum of the variables `A_vector` and `B_vector` and it assign to `total_vector`. The result should be a vector.
- Inspect the result by printing `total_vector` to the console.
- Do the same thing, but this time subtract `B_vector` from `A_vector` and assign the result to `diff_vector`.
- Finally, print `diff_vector` to the console as well.

*** =hint
Use the `+` operator to sum `A_vector` and `B_vector`. Use `<-` to assign the result to `total_vector`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# A_vector and B_vector have already been defined for you
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector: total_vector

  
# Print total_vector


# Calculate the difference between A_vector and B_vector: diff_vector


# Print diff_vector

```

*** =solution
```{r eval=FALSE}
# A_vector and B_vector have already been defined for you
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector: total_vector
total_vector <- A_vector + B_vector
  
# Print total_vector
total_vector

# Calculate the difference between A_vector and B_vector: diff_vector
diff_vector <- A_vector - B_vector

# Print diff_vector
diff_vector
```

*** =sct
```{r eval=FALSE}
test_error()
msg <- "Do not remove or change the definition of the vectors <code>A_vector</code> and <code>B_vector</code>. These have already been defined for you."
test_object("A_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("B_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("total_vector")
test_output_contains("total_vector", incorrect_msg = "Don't forget to print <code>total_vector</code> to the console.")
test_object("diff_vector")
test_output_contains("total_vector", incorrect_msg = "Don't forget to print <code>diff_vector</code> to the console.")
success_msg("Good job! Continue to the next exercise.")
```

*** =skills
1

--- type:NormalExercise xp:100
## Calculate your earnings

Now that you understand how R does arithmetic calculations with vectors, it is time to get those Ferraris in your garage! First, you need to understand what the overall profit or loss per day of the week was. The total daily profit is the sum of the profit/loss you realized on poker per day, and the profit/loss you realized on roulette per day.

*** =instructions
Assign to the variable `total_daily` how much you won or lost on each day in total (poker and roulette combined).

*** =hint
Similar to the previous exercise, assign the sum of two vectors to a new variable, `total_daily`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate your daily earnings: total_daily

```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate your daily earnings: total_daily
total_daily <- poker_vector + roulette_vector
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("total_daily", 
            incorrect_msg = "Make sure that you assign the sum of <code>poker_vector</code> and <code>roulette_vector</code> to <code>total_daily</code>.")

success_msg("Great! Continue to the next exercise.")
```

*** =skills
1

--- type:NormalExercise xp:100
## Calculate total winnings: sum()

Based on the previous analysis, it looks like you had a mix of good and bad days. This is not what your ego expected, and you wonder if there may be a (very very very) tiny chance you have lost money over the week in total? 

You can answer this question using the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function. As mentioned in the video, it calculates the sum of all elements of a vector.

*** =instructions
- Calculate the total amount of money that you have won/lost with poker and assign it to the variable `total_poker`.
- Do the same thing for roulette and assign the result to `total_roulette`.
- Now that you have the totals for roulette and poker, you can easily calculate `total_week` (which is the sum of all gains and losses of the week).
- Print the variable `total_week`.

*** =hint
Use the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function to get the total of the `poker_vector`. Do the same thing for `roulette_vector`. `total_week` is then simply `roulette_vector + poker_vector`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker: total_poker


# Total winnings with roulette: total_roulette


# Total winnings overall: total_week


# Print total_week
```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker: total_poker
total_poker <- sum(poker_vector)

# Total winnings with roulette: total_roulette
total_roulette <-  sum(roulette_vector)

# Total winnings overall: total_week
total_week <- total_roulette + total_poker

# Print total_week
total_week
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("total_poker", 
            undefined_msg = "Please make sure to define a variable <code>total_poker</code>.",
            incorrect_msg = "Make sure that you assign to <code>total_poker</code> the sum of the <code>poker_vector</code>.")
test_object("total_roulette",
            undefined_msg = "Please make sure to define a variable <code>total_roulette</code>.",
            incorrect_msg = "Make sure that you assign to <code>total_roulette</code> the sum of the <code>roulette_vector</code>.")
test_object("total_week",
            undefined_msg = "Please make sure to define a variable <code>total_week</code>.",
            incorrect_msg = "Make sure that you assign to <code>total_week</code> the sum of the other two total vectors.")
test_output_contains("total_week", incorrect_msg = "Don't forget to print the resulting variable <code>total_week</code> so that you can inspect it.")
success_msg("Oops, it seems like you are losing money. Time to rethink and adapt your strategy! This will require some deeper analysis...")
```

*** =skills
1

--- type:NormalExercise xp:100
## Comparing total winnings

The previous exercise showed that you are losing money, now what? After a short brainstorm in your hotel's jacuzzi, you realize that a possible explanation might be that your skills in roulette are not as well developed as your skills in poker. You choose to use the `>` operator to reveal this.

*** =instructions
- Create a new vector containing logicals, `poker_better`, that tells whether your poker gains exceeded your roulette results on a daily basis.
- Calculate `total_poker` and `total_roulette` as in the previous exercise.
- Using `total_poker` and `total_roulette`, Check if your total gains in poker are higher than for roulette by using a comparison. Assign the result of this comparison to the variable `choose_poker` and print it out.
What do you conclude, should you focus on roulette or on poker?

*** =hint
You partly calculated the answer to this question in the previous exercise already!
To check if 6 is larger than 5, you type `6 > 5`. This returns a variable with data type Logical (`TRUE` or `FALSE`).

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate poker_better


# Calculate total_poker and total_roulette, as before



# Calculate choose_poker


# Print choose_poker

```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate poker_better
poker_better <- poker_vector > roulette_vector

# Calculate total_poker and total_roulette, as before
total_poker <- sum(poker_vector)
total_roulette <- sum(roulette_vector)

# Calculate choose_poker
choose_poker <- total_poker > total_roulette

# Print choose_poker
choose_poker
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_better",
            incorrect_msg = "Have you calculated <code>poker_better</code> correctly? Use <code>poker_vector</code>, <code>roulette_vector</code> and the <code>></code> operator.")
test_object("total_poker", 
            undefined_msg = "Please make sure to define a variable <code>total_poker</code>.",
            incorrect_msg = "Make sure that you assign to <code>total_poker</code> the sum of the <code>poker_vector</code>.")
test_object("total_roulette",
            undefined_msg = "Please make sure to define a variable <code>total_roulette</code>.",
            incorrect_msg = "Make sure that you assign to <code>total_roulette</code> the sum of the <code>roulette_vector</code>.")
test_object("choose_poker",
            undefined_msg = "Please make sure to define a variable <code>choose_poker</code>.",
            incorrect_msg = "It looks like the <code>choose_poker</code> variable does not contain the correct answer. Make sure to read the assignment carefully.")
test_output_contains("choose_poker", incorrect_msg = "Don't forget to print <code>choose_poker</code> to the console.")
success_msg("Great! Your hunch seemed to be right. It appears that the poker game is more your cup of tea than roulette. Ready for a challenge? Head over to the next exercise!")
```

*** =skills
1

--- type:ChallengeExercise xp:200
## First steps in rational gambling

In the previous exercise, you found out that roulette is not really your forte. However, you have some vague memories from visits in Vegas where you actually excelled at this game. You plan to dig through your receipts of when you withdrew and cashed chips and found out about your actual performance in the previous week you were in Sin City. In that week, you also only played poker and roulette; the information is stored in `poker_past` and `roulette_past`. The information for the current week, with which you have been working all along, is in `poker_present` and `roulette_present`. All these variables are available in your workspace.

*** =challenge_steps
## Calculate the total gains for the past
Use the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function twice in combination with the `+` operator to calculate the total gains for your entire past week in Vegas (this means for both poker and roulette). Assign the result to `total_past`

## Calculate difference of past to present poker performance
Using the `-` operator, subtract `poker_past` from `poker_present`, to calculate `diff_poker`.

## Were you better at roulette before?
Use the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function twice and the `>` operator to see whether you were better in roulette in the past than you are in the current week. Assign the result to `roulette_worsened`.

*** =challenge_goal
## Compare your past performance to your current performance
To better compare your performance, perform the following calculations:
- Calculate the total gains for your entire past week in Vegas and assign the result to `total_past`.
- Find the daily difference in poker performance between your current week and the past week; store the result in a variable `diff_poker`.
- Create a logical that tells whether your total winnings of roulette winnings in the past were higher than your total winnings in the current week. Assign the result to `roulette_worsened`.

*** =pre_exercise_code
```{r eval=FALSE}
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
poker_present <- c(140, -50, 20, -120, 240)
names(poker_present) <- days_vector
roulette_present <- c(-24, -50, 100, -350, 10)
names(roulette_present) <- days_vector
poker_past <- c(-70, 90, 110, -120, 30)
roulette_past <- c(20, 45, 120, -230, 40)
```

*** =solution
```{r eval=FALSE}
total_past <- sum(poker_past) + sum(roulette_past)
diff_poker <- poker_present - poker_past
roulette_worsened <- sum(roulette_past) > sum(roulette_present)
```

*** =sct
```{r eval=FALSE}
test_instruction(1, {
  test_error()
  test_object("total_past")
})
test_instruction(2, {
  test_error()
  test_object("total_past")
  test_object("diff_poker")
})
test_instruction(3, {
  test_error()
  test_object("total_past")
  test_object("diff_poker")
  test_object("roulette_worsened")
})
test_instruction(4, {
  test_error()
  test_object("total_past")
  test_object("diff_poker")
  test_object("roulette_worsened")
})
success_msg("Awesome! It seems that indeed, your roulette skills have worsened if you compare to your previous week in Vegas.")
```

*** =skills
1

--- type:VideoExercise xp:50
## Vector Subsetting

*** =video_link
```{r,eval=FALSE}
//player.vimeo.com/video/138173916
```

*** =video_stream
```{r,eval=FALSE}
https://player.vimeo.com/external/138173916.hd.mp4?s=84a72eef2563eb9f94b97ba1b626815f&profile_id=119
```

*** =video_hls
```{r,eval=FALSE}
https://player.vimeo.com/external/138173916.m3u8?p=high,standard,mobile&s=d1f7079f7e71a3cd4fdc4f34046b514f
```

*** =skills
1

--- type:NormalExercise xp:100
## Selection by index (1)

After you figured that roulette is not your forte, you decide to compare the your performance at the beginning of the working week compared to the end of it. You did have a couple of Margarita cocktails at the end of the week...

To answer that question, you only want to focus on a selection of the `total_vector`. In other words, our goal is to select specific elements of the vector.

*** =instructions
- Assign the poker results of Wednesday to the variable `poker_wednesday`.
- Assign the roulette results of Friday to the variable `roulette_friday`.

*** =hint
Wednesday is the third element of `poker_vector`, and can thus be selected with `poker_vector[3]`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Poker results of Wednesday: poker_wednesday


# Roulette results of Friday: roulette_friday

```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Poker results of Wednesday: poker_wednesday
poker_wednesday <- poker_vector[3]

# Roulette results of Friday: roulette_friday
roulette_friday <- roulette_vector[5]
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_wednesday",
            incorrect_msg = "It looks like <code>poker_wednesday</code> does not contain the correct value of <code>poker_vector</code>.")
test_object("roulette_friday",
            incorrect_msg = "It looks like <code>roulette_friday</code> does not contain the correct value of <code>roulette_vector</code>.")
success_msg("Great! R also makes it possible to select multiple elements from a vector at once, remember? Put the theory to practice in the next exercise!")
```

*** =skills
1

--- type:NormalExercise xp:100
## Selection by index (2) 

How about analyzing your midweek results? 

Instead of using a single number to select a single element, you can also select multiple elements by passing a vector inside the square brackets. For example,

```
poker_vector[c(1,5)]
```

selects the first and the fifth element of `poker_vector`.


*** =instructions
- Assign the poker results of Tuesday, Wednesday and Thursday to the variable `poker_midweek`.
- Assign the roulette results of Thursday and Friday to the variable `roulette_endweek`.

*** =hint
Use the vector `c(2,3,4)` between square brackets to select the correct elements of `poker_vector`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
``` 

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Mid-week poker results: poker_midweek


# End-of-week roulette results: roulette_endweek


```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Mid-week poker results: poker_midweek
poker_midweek <- poker_vector[c(2, 3, 4)]

# End-of-week roulette results: roulette_endweek
roulette_endweek <- roulette_vector[c(4,5)]
```

*** =sct
```{r eval=FALSE}
test_error()
msg <- "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

msg <- "It looks like <code>%s</code> does not contain the correct elements from <code>%s</code>."
test_object("poker_midweek", 
            incorrect_msg = sprintf(msg, "poker_midweek", "poker_vector"))
test_object("roulette_endweek",
            incorrect_msg = sprintf(msg, "roulette_endweek", "roulette_vector"))

success_msg("Well done! Continue to the next exercise to specialize in vector selection some more!");
```

*** =skills
1

--- type:NormalExercise xp:100
## Vector selection: the good times (3) 

Now, selecting multiple successive elements of `poker_vector` with `c(2,3,4)` is not very convenient. Many statisticians are lazy people by nature, so they created an easier way to do this: `c(2,3,4)` can be abbreviated to `2:4`, which generates a vector with all natural numbers from 2 up to 4. Try it out in the console!

So, another way to find the mid-week results is `poker_vector[2:4]`. Notice how the vector `2:4` is placed between the square brackets to select element 2 up to 4. You don't have to use the [`c()`](http://www.rdocumentation.org/packages/base/functions/c) function if you're using the shortcut with the colon.

*** =instructions
- Assign to `roulette_subset` the roulette results from Tuesday to Friday inclusive by making use of `:`.
- Print the resulting variable to the console.

*** =hint
Assign a selection of `roulette_vector` to `roulette_subset` by placing `2:5` between square brackets.

*** =sample_code

```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Roulette results for Tuesday to Friday inclusive: roulette_subset


# Print roulette_subset

```

*** =solution

```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Roulette results for Tuesday to Friday inclusive: roulette_subset
roulette_subset <- roulette_vector[2:5]

# Print roulette_subset
roulette_subset
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_subset",
            incorrect_msg = "It looks like <code>roulette_selection_vector</code> does not contain the correct selection from <code>roulette_vector</code>. Make sure to to use the right indices.")
test_student_typed("2:5", not_typed_msg = "Make sure to use the combination <code>2:5</code> inside square brackets to select the elements in a concise way.")
test_output_contains("roulette_subset", incorrect_msg = "Don't forget to print out <code>roulette_subset</code> to the console.")
success_msg("Awesome! The colon operator is extremely useful and very often used in R programming, so remember it well. Have you noticed that the elements in <code>poker_vector</code> and <code>roulette_vector</code> also have names associated with them? You can also subset vectors using these names, remember?")
```

*** =skills
1

--- type:NormalExercise xp:100
## Selection by name (1)

Another way to tackle the previous exercise is by using the names of the vector elements (Monday, Tuesday, ...) instead of their numeric positions. For example, 

```
poker_vector["Monday"]
```

will select the first element of `poker_vector` since `"Monday"` is the name of that first element.

*** =instructions
- Select the fourth element, corresponding to Thursday, from `roulette_vector`. Name it `roulette_thursday`.
- Select Tuesday's poker gains using subsetting by name. Assign the result to `poker_tuesday`.

*** =hint
You can use `mean(my_vector)` to get the mean of the vector `my_vector`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select Thursday's roulette gains: roulette_thursday


# Select Tuesday's poker gains: poker_tuesday

```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select Thursday's roulette gains: roulette_thursday
roulette_thursday <- roulette_vector["Thursday"]

# Select Tuesday's poker gains: poker_tuesday
poker_tuesday <- poker_vector["Tuesday"]
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("roulette_thursday")
test_object("poker_tuesday")
success_msg("Good job! Head over to the next exercise.");
```

*** =skills
1

--- type:NormalExercise xp:100
## Selection by name (2)

Just like selecting single elements using numerics extends naturally to selecting multiple elements, you can also use a vector of names. As an example, try

```
roulette_vector[c("Monday","Wednesday")]
```

Of course you can't use the colon trick here: `"Monday":"Wednesday"` will generate an error.

*** =instructions
- Create a vector containing the poker gains for the first three days of the week; name it `poker_start`.
- Using the function [`mean()`](http://www.rdocumentation.org/packages/base/functions/mean), calculate the average poker gains during these first three days. Assign the result to a variable `avg_poker_start`.

*** =hint
You can use `mean(my_vector)` to get the mean of the vector `my_vector`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select the first three elements from poker_vector: poker_start


# Calculate the average poker gains during the first three days: avg_poker_start

```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select the first three elements from poker_vector: poker_start
poker_start <- poker_vector[c("Monday", "Tuesday", "Wednesday")]

# Calculate the average poker gains during the first three days: avg_poker_start
avg_poker_start <- mean(poker_start)
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("poker_start")
test_object("avg_poker_start",
            incorrect_msg = "Check your code to calculate <code>avg_poker_start</code> again. Use the function <code>mean()</code> in combination with <code>poker_start</code>")
success_msg("Good job! Next to subsetting vectors by index or by name, you can also use logical vectors. The next exercises will test you on this.");
```

*** =skills
1

--- type:NormalExercise xp:100
## Selection by logicals (1)

There are basically three ways to subset vectors: by using the indices, by using the names (if the vectors are named) and by using logical vectors. Filip already told you about the internals in the instructional video. As a refresher, have a look at the following statements to select elements from `poker_vector`, which are all equivalent:

```
# selection by index
poker_vector[c(1,3)]

# selection by name
poker_vector[c("Monday", "Wednesday")]

# selection by logicals
poker_vector[c(TRUE, FALSE, TRUE, FALSE, FALSE)]
```

*** =instructions
- Assign the roulette results from the first, third and fifth day to `roulette_subset`.
- Select the first three days from `poker_vector` using a vector of logicals. Assign the result to `poker_start`.

*** =hint
The logical vector to use inside square brackets for the first instruction is `c(TRUE, FALSE, TRUE, FALSE, TRUE)`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Roulette results for day 1, 3 and 5: roulette_subset

  
# Poker results for first three days: poker_start
```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Roulette relsults for day 1, 3 and 5: roulette_subset
roulette_subset <- roulette_vector[c(TRUE, FALSE, TRUE, FALSE, TRUE)]
  
# Poker results for first three days: poker_start
poker_start <- poker_vector[c(TRUE, TRUE, TRUE, FALSE, FALSE)]
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_subset")
test_object("poker_start")
success_msg("Nice one! Using logical vectors to perform subsetting might seem somewhat tedious, but its true power will become clear in the next exercise!")
```

*** =skills
1

--- type:NormalExercise xp:100
## Selection by logicals (2)

By making use of a combination of comparison operators and subsetting using logicals, you can investigate your casino performance in a more pro-active way.

The (logical) comparison operators known to R are:
- `<` for less than
- `>` for greater than
- `<=` for less than or equal to
- `>=` for greater than or equal to
- `==` for equal to each other
- `!=` not equal to each other

Experiment with these operators in the console. The result will be a logical vector, which you can use to perform subsetting! This means that instead of selecting a subset of days to investigate yourself like before, you can simply ask R to return only those days where you realized a positive return for poker.

*** =instructions
- Check if your poker winnings are positive on the different days of the week (i.e. > 0), and assign this to `selection_vector`.
- Assign the amounts that you won on the profitable days, so a vector, to the variable `poker_profits`, by using `selection_vector`.

*** =hint
- In order to check for which days your poker gains are positive, R should check for each element of `poker_vector` whether it is larger than zero. `some_vector > 0` is the way to tell R what you are after.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Create logical vector corresponding to profitable poker days: selection_vector


# Select amounts for profitable poker days: poker_profits
 
```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Create logical vector corresponding to profitable poker days: selection_vector
selection_vector <- poker_vector > 0

# Select amounts for profitable poker days: poker_profits
poker_profits <- poker_vector[selection_vector]
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("selection_vector", 
            undefined_msg = "Please make sure to define a variable <code>selection_vector</code>.",
            incorrect_msg = "It looks like <code>selection_vector</code> does not contain the correct result. Remember that R uses element wise operations for vectors.")
test_object("poker_profits",
            undefined_msg =  "Please make sure to define a variable <code>poker_profits</code>.",
            incorrect_msg =  "It looks like <code>poker_profits</code> does not contain the correct result. Remember that R uses element wise operations for vectors.")
success_msg("Great! Continue to the next exercise!")
```

*** =skills
1

--- type:NormalExercise xp:100
## Selection by logicals (3)

To fully prepare you for the challenge that's coming, you'll do a final analysis of your casino ventures. This time, you'll use your newly acquired skills to perform advanced selection on `roulette_vector`. 

Along the way, you'll need the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function. You used it before to calculate the total winnings, so an a numeric vector. However, you can also use [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) on a logical vector; it simply counts the number of vector elements that are `TRUE`.

*** =instructions
- Assign the amounts that you made on the days that you ended positively for roulette to the variable `roulette_profits`. This vector thus contains the positive winnings of `roulette_vector`. You can do this with a one-liner!
- Calculate the sum of the amounts on these profitable days; assign the result to `roulette_total_profit`.
- Find out how many roulette days were profitable, using the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function. Store the result in a variable `num_profitable_days`.

*** =hint
For the first instruction, you can use the exact same technique as in the previous exercise. However, there's no need to specify a selection vector first; you can directly place the comparison inside the square brackets!

*** =pre_exercise_code
```{r, eval=FALSE}
# no pec
```

*** =sample_code
```{r, eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select amounts for profitable roulette days: roulette_profits


# Sum of the profitable roulette days: roulette_total_profit


# Number of profitable roulette days: num_profitable_days

```

*** =solution
```{r, eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select amounts for profitable roulette days: roulette_profits
roulette_profits <- roulette_vector[roulette_vector > 0]

# Sum of the profitable roulette days: roulette_total_profit
roulette_total_profit <- sum(roulette_profits)

# Number of profitable roulette days: num_profitable_days
num_profitable_days <- sum(roulette_vector > 0)
```

*** =sct
```{r, eval=FALSE}
test_error()
msg = "Do not change anything about the definition and naming of <code>poker_vector</code> and <code>roulette_vector</code>."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_profits")
test_object("roulette_total_profit")
test_object("num_profitable_days")
success_msg("Awesome! If you inspect the variable num_profitable_days, you'll see that is equal to 2, meaning that you only had two profitable roulette days. You can conclude that roulette is not your game, right?")
```

*** =skills
1

--- type:ChallengeExercise xp:200
## Vectors: place your bets!

By now, you should have gained some insights on how your casino habits are actually working out for you. In fact, why not decide on changing your game completely? Let's dive into the world of Blackjack for once, and analyze some game outcomes here. In short, blackjack is a game where you have to ask for cards until you arrive at a sum that is as close to 21 as possible. However, if you exceed 21, you've lost. You can be greedy and go for 21, or you can be careful and settle for 16 or so. A player wins when his or her sum, or score, exceeds that of the house. 

The sums for the player's last 7 games are stored in `player`; the house's scores are contained in `house`. Both are available in the workspace. In both cases, the scores were never higher than 21.

*** =challenge_steps
## Select player score for the third game
With square brackets, select the player's score for the third game, using any of the techniques that you've learned about. Store the result in `player_third`.

## Select winning scores
Subset the `player` vector to only select the scores that exceeded the scores of `house`, so the scores that had the player win. Use subsetting in combination with the relational operator `>`. Assign the subset to the variable `winning_scores`.

## Count poor scores
Count the number of times the score inside `player` was lower than 18. This time, you should use a relational operator in combination with [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum). Save the resulting value in a new variable, `n_low_score`.

*** =challenge_goal
## Analyze the Blackjack results
- Select the player's score for the third game; store the result in `player_third`.
- Subset `player` to select only scores that led to the player winning from the house. Save the result in `winning_scores`.
- Count the number of times the player had a score lower than 18 and save this value in `n_low_score`.  

*** =pre_exercise_code
```{r eval=FALSE}
player <- c(14,17,20,21,20,18,14)
house <- c(20,15,21,20,20,17,19)
```

*** =solution
```{r eval=FALSE}
player_third <- player[3]
winning_scores <- player[player > house]
n_low_score <- sum(player < 18)
```

*** =sct
```{r eval=FALSE}
test_instruction(1, {
  test_error()
  test_object("player_third")
})
test_instruction(2, {
  test_error()
  test_object("winning_scores")
})
test_instruction(3, {
  test_error()
  test_object("n_low_score")
})
test_instruction(4, {
  test_error()
  test_object("player_third")
  test_object("winning_scores")
  test_object("n_low_score")
})
success_msg("Awesome! Given that you've solved this challenge so gracefully, you might consider adding Blackjack to your playlist on your next trip to Vegas! This challenge concludes the chapter on vectors. The next chapter will introduce you to the two-dimensional version of vectors: matrices.")
```

*** =skills
1
