--- 
title_meta  : Chapter 5
title       : Lists
description : Lists, as opposed to vectors, can hold components of different types, just like your to-do list at home or at work. This chapter will teach you how to create, name and subset these lists!
attachments : 
  slides_link: https://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/slides/ch5_slides.pdf

--- type:VideoExercise lang:r xp:50 skills:1 key:d4fe5a84ef
## Create and Name Lists

*** =video_link
//player.vimeo.com/video/138173972

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch5_1.master.m3u8


--- type:NormalExercise lang:r xp:100 skills:1 key:b14f9472dd
## Create a list

Just a quick refresher: A list in R allows you to gather a variety of objects in an ordered way. These objects can be matrices, vectors, factors, data frames, even other lists, etc. It is not even required that these objects are related to each other. You can construct a list with the [`list()`](http://www.rdocumentation.org/packages/base/functions/list) function:

```
my_list <- list(comp1, comp2 ...)
```

*** =instructions
Construct a list, named `my_list`, that contains the variables `my_vector`, `my_matrix` and `my_factor` as list components.

*** =hint
Use the [`list()`](http://www.rdocumentation.org/packages/base/functions/list) function with `my_vector`, `my_matrix` and `my_factor` as arguments separated by a comma.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Numeric vector: 1 up to 10
my_vector <- 1:10 

# Numeric matrix: 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# Factor of sizes
my_factor <- factor(c("M","S","L","L","M"), ordered = TRUE, levels = c("S","M","L"))

# Construct my_list with these different elements

```

*** =solution
```{r}
# Numeric vector: 1 up to 10
my_vector <- 1:10 

# Numeric matrix: 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# Factor of sizes
my_factor <- factor(c("M","S","L","L","M"), ordered = TRUE, levels = c("S","M","L"))

# Construct my_list with these different elements
my_list <- list(my_vector, my_matrix, my_factor)
```

*** =sct
```{r}
test_error()
msg = "Do not remove or change the definition of the variables `my_vector`, `my_matrix` or `my_factor`!"
test_object("my_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("my_matrix", undefined_msg = msg, incorrect_msg = msg)
test_object("my_factor", undefined_msg = msg, incorrect_msg = msg)
test_object("my_list", incorrect_msg = "It looks like `my_list` does not contain the correct elements. Have another look.")
success_msg("Wonderful! Your skillset is growing at a staggering pace! Head over to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:849f04f218
## Listception: lists in lists

As mentioned before, lists can also contain other lists. This works just the same as storing other types of R objects in a list. Next to the variables `my_vector`, `my_matrix` and `my_factor` from the previous exercise, now also `my_list` is predefined. Up to you to merge them in a new list; a super list!

*** =instructions
- Construct a list, named `my_super_list`, that now contains the four predefined variables listed in the sample code (in the same order).
- Print the structure of `my_super_list` with the [`str()`](http://www.rdocumentation.org/packages/utils/functions/str) function. Be sure to enter the variables in the following order: `my_ vector`, `my_ matrix`, `my_ factor`, `my_ list`

*** =hint
Just as in the previous exercise, use the [`list()`](http://www.rdocumentation.org/packages/base/functions/list) function. This time you have to add four components.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Numeric vector: 1 up to 10
my_vector <- 1:10 

# Numeric matrix: 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# Factor of sizes
my_factor <- factor(c("M","S","L","L","M"), ordered = TRUE, levels = c("S","M","L"))

# List containing vector, matrix and factor
my_list <- list(my_vector, my_matrix, my_factor)

# Construct my_super_list with the four data structures above


# Display structure of my_super_list

```

*** =solution
```{r}
# Numeric vector: 1 up to 10
my_vector <- 1:10 

# Numeric matrix: 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# Factor of sizes
my_factor <- factor(c("M","S","L","L","M"), ordered = TRUE, levels = c("S","M","L"))

# List containing vector, matrix and factor
my_list <- list(my_vector, my_matrix, my_factor)

# Construct my_super_list with the four data structures above
my_super_list <- list(my_vector, my_matrix, my_factor, my_list)

# Display structure of my_super_list
str(my_super_list)
```

*** =sct
```{r}
test_error()
msg = "Do not remove or change the definition of the variables `my_vector`, `my_matrix`, `my_factor` or `my_list`!"
test_object("my_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("my_matrix", undefined_msg = msg, incorrect_msg = msg)
test_object("my_factor", undefined_msg = msg, incorrect_msg = msg)
test_object("my_list", undefined_msg = msg, incorrect_msg = msg)
test_object("my_super_list",
            incorrect_msg = "It looks like `my_super_list` does not contain the correct elements. It's also possible that the order is not correct. Have another look.")
test_output_contains("str(my_super_list)", incorrect_msg = "Don't forget to display the structure of `my_super_list` using the [`str()`](http://www.rdocumentation.org/packages/utils/functions/str) function.")
success_msg("Nice one! Can you see from the displayed structure how the vector, matrix and ordered factor appear twice: once in the top-level list and once in the embedded list. Next!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:f37117a766
## Create a named list (1)

Well done! Let us keep this train going! To make the elements of your list clearer, you'll often want to name them: 

```
my_list <- list(name1 = your_comp1, 
                name2 = your_comp2)
``` 

If you want to name your lists after you've created them, you can use the [`names()`](http://www.rdocumentation.org/packages/base/functions/names) function as you did with vectors. The following commands are fully equivalent to the assignment above:

```
my_list <- list(your_comp1, your_comp2)
names(my_list) <- c("name1", "name2")
```

*** =instructions
- Change the code that build `my_list` by adding names to the components. Use for `my_matrix` the name `mat`, for `my_vector` the name `vec` and for `my_factor` the name `fac`.
- Print the list to the console and inspect the output.

*** =hint
The first method of assigning names to your list components is the easiest. It starts like this: 
```
my_list <- list(vec = my_vector)
```
Add the other two components in a similar fashion.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Numeric vector: 1 up to 10
my_vector <- 1:10 

# Numeric matrix: 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# Factor of sizes
my_factor <- factor(c("M","S","L","L","M"), ordered = TRUE, levels = c("S","M","L"))

# Adapt code to add names to elements in my_list
my_list <- list(my_vector, my_matrix, my_factor)

# Print my_list to the console

```

*** =solution
```{r}
# Numeric vector: 1 up to 10
my_vector <- 1:10 

# Numeric matrix: 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)

# Factor of sizes
my_factor <- factor(c("M","S","L","L","M"), ordered = TRUE, levels = c("S","M","L"))

# Adapt code to add names to elements in my_list
my_list <- list(vec = my_vector, mat = my_matrix, fac = my_factor)

# Print my_list to the console
my_list
```

*** =sct
```{r}
test_error()
msg = "Do not remove or change the definition of the variables `my_vector`, `my_matrix` or `my_factor`!"
test_object("my_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("my_matrix", undefined_msg = msg, incorrect_msg = msg)
test_object("my_factor", undefined_msg = msg, incorrect_msg = msg)
test_object("my_list",
            incorrect_msg = "It looks like `my_list` does not contain the correct elements.")
test_object("my_list", eq_condition = "equal",
            incorrect_msg = "It looks like `my_list` does not contain the correct naming for the components.");
test_output_contains("my_list", 
                     incorrect_msg = "Don't forget to print `my_list` to the console! Simply add `my_list` on a new line in the script.")
success_msg("Great! Not only do you know how to construct lists now, you can also name them; a skill that will prove most useful in practice. Continue to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:dc5f4f6f30
## Create a named list (2)

Being a huge movie fan, you decide to start storing information on good movies with the help of lists. 

Start by creating a list for the movie "The Shining". The variables `actors` and `reviews` that you'll need have already been coded in the sample code.

*** =instructions
Create the variable `shining_list`. The list contains the movie title first as "title", then the actor names as "actors", and finally the review scores factor as "reviews". Make sure to adopt the same order, and pay attention to the correct naming!

*** =hint
Let's get you started with a chunk of code:

        shining_list <- list(title = "The Shining")
        
Can you complete the rest? You still have to add `actors_vector` and `reviews_factor` with the appropriate names.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Create actors and reviews
actors_vector <- c("Jack Nicholson","Shelley Duvall","Danny Lloyd","Scatman Crothers","Barry Nelson")
reviews_factor <- factor(c("Good", "OK", "Good", "Perfect", "Bad", "Perfect", "Good"), 
                  ordered = TRUE, levels = c("Bad", "OK", "Good", "Perfect"))

# Create shining_list

```

*** =solution
```{r}
# Create actors and reviews
actors_vector <- c("Jack Nicholson","Shelley Duvall","Danny Lloyd","Scatman Crothers","Barry Nelson")
reviews_factor <- factor(c("Good", "OK", "Good", "Perfect", "Bad", "Perfect", "Good"), 
                  ordered = TRUE, levels = c("Bad", "OK", "Good", "Perfect"))

# Create the list 'shining_list'
shining_list <- list(title = "The Shining", actors = actors_vector, reviews = reviews_factor)
```

*** =sct
```{r}
test_error()
msg = "Do not remove or change the definition of the pre-set variables!"
test_object("actors_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("reviews_factor", undefined_msg = msg, incorrect_msg = msg)
test_object("shining_list",
            incorrect_msg = "It looks like `shining_list` does not contain the correct elements.")
test_object("shining_list", eq_condition = "equal",
            incorrect_msg = "It looks like `shining_list` does not contain the correct naming for the components.")
success_msg("Perfect!")
```


--- type:VideoExercise lang:r xp:50 skills:1 key:a5e5ff2680
## Subset and Extend Lists

*** =video_link
//player.vimeo.com/video/138173990

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch5_2.master.m3u8


--- type:NormalExercise lang:r xp:100 skills:1 key:1e3b5f4b0a
## Selecting elements from a list

Your list will often be built out of numerous elements and components. Therefore, getting a single element, multiple elements, or a component out of it is not always straightforward.

To select a single element from a list, for example the first element from `shining_list`, you can any one of the following commands:

```
shining_list[[1]]
shining_list[["title"]]
shining_list$title
```

If you perform selection with single square brackets, you'll end up with another list that contains the specified elements:

```
shining_list[c(2,3)]
shining_list[c(F,T,T)]
```

*** =instructions
- Select the actors from `shining_list` and assign the result to `act`.
- Create a new list containing only the title and the reviews of `shining_list`; save the new list in `sublist`.
- Display the structure of `sublist`.

*** =hint
For the first instruction you need double brackets (or the `$`), for the second one the single brackets will do.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter5.RData"))
```

*** =sample_code
```{r}
# shining_list is already defined in the workspace

# Actors from shining_list: act


# List containing title and reviews from shining_list: sublist


# Display structure of sublist

```

*** =solution
```{r}
# shining_list is already defined in the workspace

# Actors from shining_list: act
act <- shining_list[["actors"]]

# List containing title and reviews from shining_list: sublist
sublist <- shining_list[c("title", "reviews")]

# Display structure of sublist
str(sublist)
```

*** =sct
```{r}
test_error()
msg = "Do not remove or override `shining_list`."
test_object("shining_list", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("act", incorrect_msg = "Have you correctly selected the actors from `shining_list`?")
test_object("sublist", incorrect_msg = "Have you correctly selected the title and reviews from `shining_list`? Use single brackets in combination with a vector to select multiple elements.")
test_output_contains("str(sublist)", incorrect_msg = "Don't forget to display the structure of `sublist`!")

success_msg("Nice! That was still pretty easy, right? Always be aware of this difference between `[` and `[[`!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:cbda2853c3
## Chaining your selections

Besides selecting entire list elements, it's also possible that you have to access specific parts of these components themselves. It's perfectly possible to _chain your subsetting operations_ in R.

For example, with 

```
shining_list[[2]][1]
```

you select from the second component, actors (`shining_list[[2]]`), the first element (`[1]`). When you type this in the console, you will see the answer is Jack Nicholson.

*** =instructions
- Select from the `shining_list` the last actor and assign the result to `last_actor`.
- Select from the `shining_list` the second review score (which is a factor). Store the result in `second_review`.

*** =hint
- If you want to do things nicely: `length(shining_list$actors)` gives you the number of actors, and thus the element to select.
- You can select the information of the second review with `shining_list$reviews[2, ]`.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter5.RData"))
```

*** =sample_code
```{r}
# shining_list is already defined in the workspace

# Select the last actor: last_actor


# Select the second review: second_review

```

*** =solution
```{r}
# shining_list is already defined in the workspace

# Select the last actor: last_actor
last_actor <- shining_list$actors[length(shining_list$actors)]

# Select the second review: second_review
second_review <- shining_list$reviews[2]
```

*** =sct
```{r}
test_error()
msg = "Do not remove or override `shining_list`."
test_object("shining_list", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("last_actor", 
            incorrect_msg = "Looks like `last_actor` does not equal the last actor.")
test_object("second_review",
            incorrect_msg = "It looks like `second_review` does not contain the factor that corresponds to the second review.")
success_msg("Great! Selecting elements from lists is rather easy isn't it? Continue to the next exercise.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:35ba0d6f0d
## Extending a list

You already know that the `$` as well as `[[` can be used to select elements from lists. They can also be used to extend lists. To extend `shining_list` with some personal opinion, you could do one of the following things:

```
shining_list$my_opinion <- "Love it!"
shining_list[["my_opinion"]] <- "Love it!"
```

Being proud of your first list, you shared it with the members of your movie hobby club. However, one of the senior members, a guy named M. McDowell, noted that you forgot to add the release year (1980). Given your ambitions to become next year's president of the club, you decide to add this information to the list. To fully make up for your mistake, you also decide to add the name of the director (Stanley Kubrick). 

*** =instructions
- Add the release year as a numeric to `shining_list` under the name `year`.
- Add the director to the list, `"Stanley Kubrick"`, with the name `director`.
- Finally, inspect the structure of `shining_list`.

*** =hint
Let the examples in the assignment guide you to list extension mastery! Remember that R is case sensitive!

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/chapter5.RData"))
```

*** =sample_code

```{r}
# shining_list is already defined in the workspace

# Add the release year to shining_list


# Add the director to shining_list


# Inspect the structure of shining_list

```

*** =solution
```{r}
# shining_list is already defined in the workspace

# Add the release year to shining_list
shining_list$year <- 1980

# Add the director to shining_list
shining_list$director <- "Stanley Kubrick"

# Inspect the structure of shining_list
str(shining_list)
```

*** =sct
```{r}
test_error()
test_object("shining_list", 
            incorrect_msg = paste("Have you correctly added both the year and the director to `shining_list`?",
                                  "Make sure to use the correct names (\"year\" and \"director\"). Remember that R is case sensitive!"))

test_output_contains("str(shining_list)", 
                     incorrect_msg = "Do not forget to inspect the structure of `shining_list` using the [`str()`](http://www.rdocumentation.org/packages/utils/functions/str) function.")

success_msg("Congratulations on finishing up on this chapter!")
```
