--- 
title_meta  : Chapter 2
title       : Vectors
description : We take you on a trip to Vegas, where you will learn how to analyze your gambling results using vectors in R! After completing this chapter, you will be able to create vectors in R, name them, select elements from them and compare different vectors.
attachments : 
  slides_link: https://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/slides/ch2_slides.pdf
  
--- type:VideoExercise lang:r xp:50 skills:1 key:b91dd847a0
## Create and Name Vectors

*** =video_link
//player.vimeo.com/video/138173896

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch2_1.master.m3u8


--- type:NormalExercise lang:r xp:100 skills:1 key:2d1cb04427
## Create a vector (1)

Feeling lucky? You better, because we'll take you on a trip to Las Vegas!

Thanks to R and your new data science skills, you will learn how to uplift your performance at the tables and fire off your career as a professional gambler. This chapter will show how you can easily keep track of your betting progress and how you can do some simple analyses on past actions.

You will use vectors. As Filip explained you, vectors are one dimensional arrays that can hold numeric data, character data or logical data. You create a vector with the combine function [`c()`](http://www.rdocumentation.org/packages/base/functions/c). You place the vector elements separated by a comma between the brackets. For example:

```
numeric_vector <- c(1, 2, 3)
character_vector <- c("a", "b", "c")
logical_vector <- c(TRUE, FALSE)
```

*** =instructions 
Create a vector, `logical_vector`, that contains the three elements: `TRUE`, `FALSE` and `TRUE` (in that order). 

*** =hint 
Assign `c(TRUE, FALSE, TRUE)` to the variable `logical_vector` with the `<-` operator.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("x", "y", "z")

# Create logical_vector

```

*** =solution
```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("x", "y", "z")

# Create logical_vector
logical_vector <- c(TRUE, FALSE, TRUE)
```

*** =sct
```{r}
msg <- "Do not change how `numeric_vector` and `character_vector` are created!"
lapply(c("numeric_vector", "character_vector"), test_object, undefined_msg = msg, incorrect_msg = msg)
test_object("logical_vector", incorrect_msg = "Make sure that you assign the correct values to `logical_vector`. The order matters!")
success_msg("Perfect! Let's practice some more with vector creation.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:c6e056b9c3
## Create a vector (2)

After one week in Las Vegas and still zero Ferraris in your garage, you decide that it is time to start using your data science superpowers.

Before doing your first analysis, you decide to collect all the winnings and losses for the last week: 

For `poker_vector`: 
- On Monday you won \$140
- Tuesday you **lost** \$50
- Wednesday you won \$20 
- Thursday you **lost** \$120
- Friday you won \$240

For `roulette_vector`: 
- On Monday you **lost** \$24
- Tuesday you **lost** \$50
- Wednesday you won \$100
- Thursday you **lost** \$350
- Friday you won \$10

To be able to use this data in R, you decide to create the variables `poker_vector` and `roulette_vector`.

*** =instructions
Assign the winnings/losses for roulette as a vector to the variable `roulette_vector`. Make sure to use the correct order.

*** =hint
To help you with this step, the script already contains the code for creating `poker_vector`. Assign the correct values to `roulette_vector` based on the numbers in the assignment. Do not forget that losses are negative numbers.


*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday: roulette_vector

```

*** =solution

```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday: roulette_vector
roulette_vector <- c(-24, -50, 100, -350, 10)
```

*** =sct
```{r}
test_object("poker_vector", 
            incorrect_msg = "Don't change how `poker_vector` is defined.")
test_object("roulette_vector", 
            incorrect_msg = paste("Make sure that you assign a vector with the correct values to `roulette_vector`.",
                                  "If you lost money, you should use a `-` sign."))
success_msg("Very good! To check out the contents of your vectors, remember that you can always simply type the variable in the console and hit Enter. Proceed to the next exercise!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ebb5aae2ff
## Naming a vector (1)

As a data analyst, it is important to have a clear view on the data that you are using. Understanding what each element refers to is essential. 

In the previous exercise, we created a vector with your winnings over the week. Each vector element refers to a day of the week but it is hard to tell which element belongs to which day. It would be nice if you could show that in the vector itself. Remember the [`names()`](http://www.rdocumentation.org/packages/base/functions/names) function to name the elements of a vector?

```
some_vector <- c("Johnny", "Poker Player")
names(some_vector) <- c("Name", "Profession")
```

*** =instructions
`poker_vector` has already been named with the days of the week. Do the same thing for `roulette_vector`. Beware: R is case sensitive!

*** =hint
Assign `c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")` to `names(roulette_vector)`.

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Add names to poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Add names to roulette_vector

```

*** =solution
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Add names to poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Add names to roulette_vector
names(roulette_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```

*** =sct
```{r}
msg <- "Do not change the values inside `%s`; they were already coded for you."
test_object("poker_vector", incorrect_msg = sprintf(msg, "poker_vector"))
test_object("roulette_vector", incorrect_msg = sprintf(msg, "roulette_vector"))
msg <- "Make sure that you assign the correct names vector to `%s`. The names of the day should start with a capital letter!"
test_object("poker_vector", eq_condition = "equal", incorrect_msg = sprintf(msg, "poker_vector"))
test_object("roulette_vector", eq_condition = "equal", incorrect_msg = sprintf(msg, "roulette_vector"))
success_msg("Well done!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:5c026ed9fb
## Naming a vector (2)

If you want to become a good statistician, you have to become lazy. (If you are already lazy, chances are high you are one of those exceptional, natural-born statistical talents!)

In the previous exercises you probably experienced that it is boring and frustrating to type and retype information such as the days of the week. However, there is a more efficient way to do this, namely, to assign the days of the week vector to a variable! 

Just like you did with your poker and roulette returns, you can also create a variable that contains the days of the week. This way you can use and re-use it. This variable, `days_vector`, has already been coded for you.

*** =instructions
- Use the variable `days_vector` to set the names of `poker_vector`.
- Use the variable `days_vector` to set the names of `roulette_vector`.

*** =hint
You can use `names(poker_vector) <- ` to set the names of the variable `poker_vector`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Create the variable days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
 
# Use days_vector to name poker_vector


# Use days_vector to name roulette_vector
```

*** =solution
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Create the variable days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Use days_vector to name poker_vector
names(poker_vector) <- days_vector

# Use days_vector to name roulette_vector
names(roulette_vector) <- days_vector
```

*** =sct
```{r}
msg <- "Do not change the values inside `%s`; they were already coded for you."
test_object("poker_vector", incorrect_msg = sprintf(msg, "poker_vector"))
test_object("roulette_vector", incorrect_msg = sprintf(msg, "roulette_vector"))
test_object("days_vector", incorrect_msg = sprintf(msg, "days_vector"))

msg <- "Make sure that you assign `days_vector` to the names of `%s`. Use the `names()` function."
test_object("poker_vector", eq_condition = "equal", incorrect_msg = sprintf(msg, "poker_vector"))
test_object("roulette_vector", eq_condition = "equal", incorrect_msg = sprintf(msg, "roulette_vector"))

success_msg("Nice one! A word of advice: try to avoid code duplication at all times.")
```

--- type:VideoExercise lang:r xp:50 skills:1 key:b47466f033
## Vector Arithmetic

*** =video_link
//player.vimeo.com/video/141163398

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch2_2.master.m3u8


--- type:NormalExercise lang:r xp:100 skills:1 key:6b17fc50b9
## Calculate your earnings

Now that you understand how R does arithmetic calculations with vectors, it is time to get those Ferraris in your garage! First, you need to understand what the overall profit or loss per day of the week was. The total daily profit is the sum of the profit/loss you realized on poker per day, and the profit/loss you realized on roulette per day.

Remember that vector calculations happen element-wise; the following three statements are completely equivalent:

```
c(1, 2, 3) + c(4, 5, 6)
c(1 + 4, 2 + 5, 3 + 6)
c(5, 7, 9)
```

*** =instructions
- Assign to the variable `total_daily` how much you won or lost on each day in total (poker and roulette combined). `total_daily` should be a vector with 5 values.
- Print out `total_daily`.

*** =hint
Similar to the previous exercise, assign the sum of two vectors to a new variable, `total_daily`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate your daily earnings: total_daily


# Print out total_daily
```

*** =solution
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate your daily earnings: total_daily
total_daily <- poker_vector + roulette_vector

# Print out total_daily
total_daily
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("total_daily", 
            incorrect_msg = "Make sure that you assign the sum of `poker_vector` and `roulette_vector` to `total_daily`. Simply use `+`.")
test_output_contains("total_daily", incorrect_msg = "Don't forget to print out `total_daily`.")
success_msg("Great! Continue to the next exercise.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:a9a1a50a31
## Calculate total winnings: sum()

Based on the previous analysis, it looks like you had a mix of good and bad days. This is not what your ego expected, and you wonder if there may be a (very very very) tiny chance you have lost money over the week in total? 

You can answer this question using the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function. As mentioned in the video, it calculates the sum of all elements of a vector.

*** =instructions
- Calculate the total amount of money that you have won/lost with poker and assign it to the variable `total_poker`.
- Do the same thing for roulette and assign the result to `total_roulette`.
- Use `+` to sum the `total_poker` and `total_roulette`, which is the sum of all gains and losses of the week. Simply print the result to the console.

*** =hint
Use the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function to get the total of the `poker_vector`. Do the same thing for `roulette_vector`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker: total_poker


# Total winnings with roulette: total_roulette


# Total winnings overall: print out the result

```

*** =solution
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker: total_poker
total_poker <- sum(poker_vector)

# Total winnings with roulette: total_roulette
total_roulette <-  sum(roulette_vector)

# Total winnings overall: print out the result
total_roulette + total_poker
```

*** =sct
```{r}
msg <- "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("total_poker", 
            undefined_msg = "Please make sure to define a variable `total_poker`.",
            incorrect_msg = "Make sure that you assign to `total_poker` the sum of the `poker_vector`.")
test_object("total_roulette",
            undefined_msg = "Please make sure to define a variable `total_roulette`.",
            incorrect_msg = "Make sure that you assign to `total_roulette` the sum of the `roulette_vector`.")
test_output_contains("total_poker + total_roulette", incorrect_msg = "Print the sum of `total_poker` and `total_roulette` to the console.")
success_msg("Oops, it seems like you are losing money. Time to rethink and adapt your strategy! This will require some deeper analysis...")
```


--- type:VideoExercise lang:r xp:50 skills:1 key:513029f4ac
## Vector Subsetting

*** =video_link
//player.vimeo.com/video/138173916

*** =video_hls
//videos.datacamp.com/transcoded/732_intro_to_r/v1/hls-ch2_3.master.m3u8


--- type:NormalExercise lang:r xp:100 skills:1 key:6112e74425
## Selection by index (1)

After you figured that roulette is not your forte, you decide to compare your performance at the beginning of the week to your performance at the end of the week. You did have a couple of Margarita cocktails at the end of the week...

To answer that question, you only want to focus on a selection of the `total_vector`. In other words, our goal is to select specific elements of the vector.

*** =instructions
- Assign the poker results of Wednesday to the variable `poker_wednesday`.
- Assign the roulette results of Friday to the variable `roulette_friday`.

*** =hint
Wednesday is the third element of `poker_vector`, and can thus be selected with `poker_vector[3]`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Poker results of Wednesday: poker_wednesday


# Roulette results of Friday: roulette_friday

```

*** =solution
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Poker results of Wednesday: poker_wednesday
poker_wednesday <- poker_vector[3]

# Roulette results of Friday: roulette_friday
roulette_friday <- roulette_vector[5]
```

*** =sct
```{r}

msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_wednesday",
            incorrect_msg = "It looks like `poker_wednesday` does not contain the correct value of `poker_vector`.")
test_object("roulette_friday",
            incorrect_msg = "It looks like `roulette_friday` does not contain the correct value of `roulette_vector`.")
success_msg("Great! R also makes it possible to select multiple elements from a vector at once, remember? Put the theory to practice in the next exercise!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:ae2832fbd1
## Selection by index (2) 

How about analyzing your midweek results? 

Instead of using a single number to select a single element, you can also select multiple elements by passing a vector inside the square brackets. For example,

```
poker_vector[c(1,5)]
```

selects the first and the fifth element of `poker_vector`.


*** =instructions
- Assign the poker results of Tuesday, Wednesday and Thursday to the variable `poker_midweek`.
- Assign the roulette results of Thursday and Friday to the variable `roulette_endweek`.

*** =hint
Use the vector `c(2,3,4)` between square brackets to select the correct elements of `poker_vector`.

*** =pre_exercise_code
```{r}
# no pec
``` 

*** =sample_code
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Mid-week poker results: poker_midweek


# End-of-week roulette results: roulette_endweek


```

*** =solution
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Mid-week poker results: poker_midweek
poker_midweek <- poker_vector[c(2, 3, 4)]

# End-of-week roulette results: roulette_endweek
roulette_endweek <- roulette_vector[c(4,5)]
```

*** =sct
```{r}
msg <- "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

msg <- "It looks like `%s` does not contain the correct elements from `%s`."
test_object("poker_midweek", 
            incorrect_msg = sprintf(msg, "poker_midweek", "poker_vector"))
test_object("roulette_endweek",
            incorrect_msg = sprintf(msg, "roulette_endweek", "roulette_vector"))

success_msg("Well done! Another way to find the mid-week results is `poker_vector[2:4]`. Continue to the next exercise to specialize in vector selection some more!");
```

--- type:NormalExercise lang:r xp:100 skills:1 key:5919f3fc05
## Selection by name

Another way to tackle the previous exercise is by using the names of the vector elements (Monday, Tuesday, ...) instead of their numeric positions. For example, 

```
poker_vector["Monday"]
```

will select the first element of `poker_vector` since `"Monday"` is the name of that first element.

*** =instructions
- Select the fourth element, corresponding to Thursday, from `roulette_vector`. Name it `roulette_thursday`.
- Select Tuesday's poker gains using subsetting by name. Assign the result to `poker_tuesday`.

*** =hint
You can use `mean(my_vector)` to get the mean of the vector `my_vector`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select Thursday's roulette gains: roulette_thursday


# Select Tuesday's poker gains: poker_tuesday

```

*** =solution
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select Thursday's roulette gains: roulette_thursday
roulette_thursday <- roulette_vector["Thursday"]

# Select Tuesday's poker gains: poker_tuesday
poker_tuesday <- poker_vector["Tuesday"]
```

*** =sct
```{r}
msg <- "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("roulette_thursday")
test_object("poker_tuesday")
success_msg("Good job! Head over to the next exercise.");
```

--- type:NormalExercise lang:r xp:100 skills:1 key:22121c6c46
## Selection by logicals (1)

There are basically three ways to subset vectors: by using the indices, by using the names (if the vectors are named) and by using logical vectors. Filip already told you about the internals in the instructional video. As a refresher, have a look at the following statements to select elements from `poker_vector`, which are all equivalent:

```
# selection by index
poker_vector[c(1,3)]

# selection by name
poker_vector[c("Monday", "Wednesday")]

# selection by logicals
poker_vector[c(TRUE, FALSE, TRUE, FALSE, FALSE)]
```

*** =instructions
- Assign the roulette results from the first, third and fifth day to `roulette_subset`.
- Select the first three days from `poker_vector` using a vector of logicals. Assign the result to `poker_start`.

*** =hint
The logical vector to use inside square brackets for the first instruction is `c(TRUE, FALSE, TRUE, FALSE, TRUE)`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Roulette results for day 1, 3 and 5: roulette_subset

  
# Poker results for first three days: poker_start
```

*** =solution
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Roulette relsults for day 1, 3 and 5: roulette_subset
roulette_subset <- roulette_vector[c(TRUE, FALSE, TRUE, FALSE, TRUE)]
  
# Poker results for first three days: poker_start
poker_start <- poker_vector[c(TRUE, TRUE, TRUE, FALSE, FALSE)]
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_subset")
test_object("poker_start")
success_msg("Nice one! Using logical vectors to perform subsetting might seem somewhat tedious, but its true power will become clear in the next exercise!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:aa2e5f6e97
## Selection by logicals (2)

By making use of a combination of comparison operators and subsetting using logicals, you can investigate your casino performance in a more pro-active way.

The (logical) comparison operators known to R are:
- `<` for less than
- `>` for greater than
- `<=` for less than or equal to
- `>=` for greater than or equal to
- `==` for equal to each other
- `!=` not equal to each other

Experiment with these operators in the console:

```
lost_roulette_days <- roulette_vector < 0
lost_roulette_days
```

The result will be a logical vector, which you can use to perform subsetting, like this example:

```
roulette_vector[lost_roulette_days]
```

The result is a subset of `roulette_vector` that contains only your losses in roulette.

*** =instructions
- Check if your poker winnings are positive on the different days of the week (i.e. > 0), and assign this to `selection_vector`.
- Assign the amounts that you won on the profitable days to the variable `poker_profits` by using `selection_vector`.

*** =hint
- In order to check for which days your poker gains are positive, R should check for each element of `poker_vector` whether it is larger than zero. `some_vector > 0` is the way to tell R what you are after.
- After creating `selection_vector`, you can use it to subset `poker_vector` like this: `poker_vector[selection_vector]`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Create logical vector corresponding to profitable poker days: selection_vector


# Select amounts for profitable poker days: poker_profits
 
```

*** =solution
```{r}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Create logical vector corresponding to profitable poker days: selection_vector
selection_vector <- poker_vector > 0

# Select amounts for profitable poker days: poker_profits
poker_profits <- poker_vector[selection_vector]
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("selection_vector", 
            undefined_msg = "Please make sure to define a variable `selection_vector`.",
            incorrect_msg = "It looks like `selection_vector` does not contain the correct result. Remember that R uses element wise operations for vectors.")
test_object("poker_profits",
            undefined_msg =  "Please make sure to define a variable `poker_profits`.",
            incorrect_msg =  "It looks like `poker_profits` does not contain the correct result. Remember that R uses element wise operations for vectors.")
success_msg("Great! Move on to the Matrices chapter!")
```

