
## Lists

A list is a generic vector containing other objects. There is no particular need for the objects to be of the same type, as with vectors. For example, a list could consist of a numeric vector, a logical value, a matrix, other lists, and so on.

```{r}
my_family <- list("Ryan", "Mary", 3, TRUE)
my_family
``` 

Components of lists may also be named. You can assign names to list elements by the [`names()`](http://www.rdocumentation.org/packages/base/functions/names) function or at the time of creation.

```{r}
my_family <- list(father="Ryan", mother="Mary", siblings=3, divorced=TRUE)
my_family
```

If you want to know if your object,`my_family` in our case, is a list you can use the following. 

```{r}
is.list(my_family)
```

Finally, you can use `str()` to display the stucture of your list. 

```{r}
str(my_family)
```


## Subset and Extend Lists

If you need to isolate parts of your list you can use `[...]` and `[[...]]`. Indexing with `[...]` as used to subset vectors will give you sublist not the content inside the element. To retrieve the content, we need to use `[[...]]`. This approach will allow you to access a single element at a time.

```{r}
my_family[1]
my_family[[1]]
```

If you want to retrieve more elements of your list. 

```{r}
my_family[c(1,3)]
```

If the list is named its elements can be refered by names instead of numeric indeces.

```{r}
my_family[["father"]]
```

Alternatively, you can use the `$` operator. 
```{r}
my_family$father
```

Another way to subset is logical data types, namely `TRUE` and `FALSE`.

```{r}
my_family[c(TRUE, FALSE, TRUE, FALSE)]
```

Adding new elements is easy. You simply assign values using new tags and it will pop into action. 

```{r}
grandparents <- c("Arthur","Josephin")
my_family$grandparents <- grandparents
my_family
```

or equivalently
```{r}
my_family[["grandparents"]] <- grandparents
```



