---
output: html_document
---

## R: The true basics  


### R as calculator

After R is started, there is a console awaiting for input. At the prompt (`>`), you can enter numbers and perform calculations. 

```{r}
 3*2
```
A few arithmetic operators are:  

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`
- Modulo: `%%`


### Variable assignment and Operations

You can assign values to variables with the assignment operator `<-`. Just typing the variable by itself at the prompt will print out the value. 

```{r}
 x <- 3
 x
 y <- 9
 y
```

You can also perform arithmetic operations with variables. Look at the result of multiplying `x` and `y`, we defined previously:

```{r}
 y * x
```

As you work in R and create new variables it can be easy to lose track of what variables you have defined. To get a list of all the variables that have been defined use [`ls()`](http://www.rdocumentation.org/packages/base/functions/ls). And if you need to remove variables, you can use `rm(<var_name>)`. 


### Comment your code

Adding comments to your code helps others understanding it. Comments in R are ignored by the parser. Any text that could be typed after the `#` character and on the same line is taken to be a comment, unless the `#` character is inside a quoted string. For example,

```{r}
 x <- 24 # this is a comment 
 y <- "  #... but this is not."
```


## Basic Data Types 

There are several basic R data types that are of frequent occurrence in routine R calculations. We will try to understand a few of them better by using the [`class()`](http://www.rdocumentation.org/packages/base/functions/class) function. 

- Decimal values are called numerics in R. You can perform arithmetic operations on them.
```{r}
 x <- 12.3
 x
 class(x)
```

- A special type of numeric is an integer. You can specify that a number is an integer using the following syntax.
```{r}
 y <- 3L
 y
 class(y)
```

Another way is to invoke the [`as.integer()`](http://www.rdocumentation.org/packages/base/functions/integer) function. 

```{r}
 y <- as.integer(3)
 y
 class(y)
```

And one can convert an integer value to a numeric value by [`as.numeric()`](http://www.rdocumentation.org/packages/base/functions/numeric).  

- A character object is used to represent string values in R.

```{r}
 z <- "Good morning!"
 z
 class(z)
```

You can convert objects into character values with the [`as.character()`](http://www.rdocumentation.org/packages/base/functions/numeric) function. 

- Another important data type is the logical type. There are two predefined variables, `TRUE` and `FALSE`. 

```{r}
 a <- TRUE
 a
 class(a)
```

You can also see the data type of a variable by invoking one of the following `is.*()` functions. The result is a logical statement `TRUE` or `FALSE`. 

```{r, eval=FALSE}
 is.numeric()      #to evaluate if type = numeric
 is.integer()      #to evaluate if type = integer
 is.character()    #to evaluate if type = character
```





