## Create and Name Matrices

Matrices are not very different from vectors; both are data structures that store elements of the same type:

A matrix is a 2-dimensional array consisting of rows and columns. Matrices and dataframes are different since the latter can only contain numeric vectors and can be considered as a natural extension of a vector.
You can build them easily with the function [`matrix()`](http://www.rdocumentation.org/packages/gmp/functions/matrix).

```{r, eval= F }
matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE, dimnames = NULL) 
```

Only the number of rows `nrow` and columns `ncol` need to be specified. However, the argument `byrow` can be used to specify whether the matrix is filled up row-wise or column-wise.

```{r}
my_matrix <- matrix(c(9,2,5, 1,3,4, 1,2,7), nrow = 3, ncol = 3, byrow = TRUE)
my_matrix
```


Per default, the rows and columns do not have names. The argument `dimnames` can change that by defining a list with names such as `dimnames = list(c(r1, r2...), c(c1, c2,...)` depending on the number of rows and columns.

Since matrices are just several vectors that you can put together, they can also be build by pasting rows or columns with [`rbind()`](http://www.rdocumentation.org/packages/dplyr/functions/rbind) or [`cbind()`](http://www.rdocumentation.org/packages/marray/functions/cbind) instead of using the function [`matrix()`](http://www.rdocumentation.org/packages/gmp/functions/matrix).

A matrix is defined as an atomic vector. Thus, it is possible to create a matrix based on two matrices that do not necessary contain numbers, as seen in the first video exercise. Then you have created a dataframe or a list by applying coercion.


## Subsetting Matrices
> Like any other data object, you can draw subsets from the matrices. They can be built using the square brackets `[]` on the matrix object and specifying the row and column that is to be subtracted. You can then maintain a single matrix element:

```{r}
my_matrix[1, 2]
```

Otherwise rows and columns can be specifiyed by simply defining the number of the row or the column:

```{r}
my_matrix[1,]
my_matrix[, 2]
```

If only a single number is defined, R returns the value of the position defined inside of the subset:

```{r}
my_matrix[1]
```

Remark: R counts the positions inside of a matrix from the first row value to the last row value in the first column, then the first row value to the last row value in the second column. 

Furthermore, you can subset multiple elements of a matrix vector by defining the row or columns and the position of the value. As seen in the video lectures you can use the concatenate function [`c()`](http://www.rdocumentation.org/packages/onion/functions/c) to either retain a single value of a specific position or a sub-matrix.

```{r}
my_matrix[2, c(2, 3)]
my_matrix[c(2, 3), c(2, 3)]
```

In a similar manner, matrices can be subsetted by names instead of indices of rows and columns.
Alternatively, logical vectors can be used to subset, when both rows and columns are defined!

```{r}
my_matrix[c(F, F, T), c(F, F, T)]
```

## Matrix calculus

R has two easy functions to let you sum up the values of the rows and columns:

* [`rowSums()`](http://www.rdocumentation.org/packages/base/functions/colSums) 
* [`colSums()`](http://www.rdocumentation.org/packages/base/functions/colSums)

And of course any arithmetic operation can be proceeded on a matrix as well:

* calculate a scalar 
* any other operations (`/`, `+`, `-`) 

In general, all matrix operations are done element-wise.

Remark: Matrix recycling is automatically done when a matrix calculation is done between two unequal matrices or between a matrix and a vector. This has to be handled very carefully, since R might recylces in a way you don't want it to recycle.






