---
output: html_document
---
## Factors

Factors are defined as categorical variables that take on a few values. To define a variable as categorical use the function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor).

What does R do?

* screening for all values and defining them as factors.
* sort them alphabetically
* character values correspond to integer values (handy, in the case of long charcater strings)

## Rename factors

Moreover, the order has to be specified manually inside of the factor using the `levels` argument in the  [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) function.

```{r, eval=FALSE}
factor(my_var, levels = c("xy","xz","zy"))
```

And the level names can be defined manually using the [`levels()`](http://www.rdocumentation.org/packages/base/functions/levels) function

```{r, eval=FALSE}
levels(my_var) = c("na_xy", "na_xz", "na_zy")
```

or by using the `labels` argument inside of the function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor)

```{r, eval=F}
factor(my_var, labels = c("na_xy","na_xz","na_zy"))

```

Remark: To rename levels, you always have to follow the original order of the levels. To avoid confusion and misspecification, it is suggested to use both `levels` and `labels` inside [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor).


## Nominal vs Ordinal

**Ordinal** variables contain a natural order among their levels, whereas **nominal** variables do not inherit any such order.

First thing to know about **ordinal** variables is that they are as well defined with [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) but the argument `order` is specified as `TRUE`.
R orders them alphabetically, unless specified otherwise.
The ordinal structure is quite specific:
* it is regarded in comparisons and operations
* it reflects by `<` and `>` signs

For example:
```{r, eval=F}
factor(my_ordinal, order = T, levels = c(1, 2, 3))
```




