
## Basic Graphics

One of the most frequently used plotting functions in R is the [`plot()`](http://www.rdocumentation.org/packages/graphics/functions/plot). This is a generic function: the type of plot produced is dependent on the type or class of the argument(s).

```{r, eval=FALSE}
x <- c(1, 2, 3, 4)
plot(x) # this generates a plot of the values in the variable against their index

x <- factor(c("Black", "White", "Green"))
plot(x) # this generates a bar chart
```

```{r, eval=FALSE}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
plot(x, y) # this generates a scatter plot

x <- factor(c("Black", "White", "Green"))
y <- c(1, 2, 3)
plot(x, y) # this generates boxplots of y for each level of x

x <- factor(c("Black", "White", "Green"))
y <- factor(c("Left", "Right", "Centre"))
plot(x, y) # this generates stacked bar chart 
```

Histograms can be created using the [`hist()`](http://www.rdocumentation.org/packages/graphics/functions/hist) function. This function takes in a continuous variable, `x`, for which the histogram is plotted. 

```{r, eval=FALSE}
hist(x, breaks = ``)
```

With the `breaks` argument you can specify the number of bins you want in the histogram.

You can also check other graphics functions such as [`boxplot()`](http://www.rdocumentation.org/packages/graphics/functions/boxplot) and [`barplot()`]( http://www.rdocumentation.org/packages/raster/functions/barplot). 


## Customizing Plots

Now, you can modify your plot !

```{r,eval=FALSE}
plot(x,y, 
      xlab = " ",     # changes the label of the horizontal axis 
      ylab = " ",     # changes the label of the vertical axis 
      main = " ",     # specifies the title of the plot 
      type = " ",     # specifies the type of the plot i.e lines, points etc 
      col = " ")      # specifies the color of the plot 
```

Type `?par` in your console to take a peek on the graphical parameters you can specify. 
 
A few of them are 

```{r,eval=FALSE}
plot(x,y, 
      xlab = " ",   
      ylab = " ",  
      main = " ",   
      type = " ",   
      col = " ",
      col.main = " ",  # specifies the color of the main title
      cex.axis = ` `,  # specifies the size of the fonts 
      lty = ` `,       # specifies the line type
      pch = ` `)       # specifies the plot symbol
```

### Important Note
Since all the arguments are specified inside the [`plot()`](http://www.rdocumentation.org/packages/graphics/functions/plot)
function they are valid only for the specific plot. It is possible, though, to set the parameters of the graphs globally by using the [`par()`](http://www.rdocumentation.org/packages/graphics/functions/par) function. 


## Multiple Plots

R makes it easy to combine multiple plots into one overall graph, using either the [`par()`](http://www.rdocumentation.org/packages/graphics/functions/par)
or [`layout()`](http://www.rdocumentation.org/packages/graphics/functions/layout) function.

With [`par()`](http://www.rdocumentation.org/packages/graphics/functions/par), you can include the option `mfrow` to create a grid of `nrows` and `ncols` plots that are filled in by row. 

```{r,eval=FALSE}
par(mfrow = c(nrows, ncols))
```

If you use `mfcol`, it fills in the grid by columns.

```{r,eval=FALSE}
par(mfcol = c(nrows, ncols))
```

In order to reset the graphical parameters you can use: 

```{r,eval=FALSE}
par(mfrow = c(1, 1))
```

or equivalently,

```{r,eval=FALSE}
old_par <- par()
```

and invoke the `old_par` when you need to reset the parameters. 

Another way is to use the [`layout()`](http://www.rdocumentation.org/packages/graphics/functions/layout) function which divides the plotting space into as many rows and columns as there are in matrix `mat`.

```{r,eval=FALSE}
layout(mat, ...)
```

Once more way to reset the graphical parameters, is to use:

```{r,eval=FALSE}
layout(1)
```


In order to add more information to your plot you can use the following fuunctions. 

```{r,eval=FALSE}
plot(x, y)
abline()      # adds one or more straight lines
lines()       # adds lines (careful how to specify the arguments, watch video for more info)
points()      # adds points 
text()        # adds text 
segments()    # adds line segments between pairs of points
```

Take a look at the documentatin to get more insight into these functions [`abline()`](http://www.rdocumentation.org/packages/graphics/functions/abline),
[`lines()`](http://www.rdocumentation.org/packages/graphics/functions/lines),
[`points()`](http://www.rdocumentation.org/packages/graphics/functions/points),
[`text()`](http://www.rdocumentation.org/packages/graphics/functions/text),
and [`segments()`](http://www.rdocumentation.org/packages/graphics/functions/segments).



