--- type:MultipleChoiceExercise xp:100
## Different ways to create and name vectors

The previous exercises outlined different ways of creating and naming vectors. Have a look at this chunk of code:

```
poker_vector1 <- c(140, -50, 20, -120, 240)
names(poker_vector1) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

poker_vector2 <- c(Monday = 140, -50, 20, -120, 240)

roulette_vector1 <- c(-24, -50, 100, -350, 10)
days_vector <- names(poker_vector1)
names(roulette_vector1) <- days_vector

roulette_vector2 <- c(-24, -50, 100, -350, 10)
names(roulette_vector2) <- "Monday"
```

Which of the following statements is true?

*** =instructions
- The code to define `poker_vector2` is syntactically invalid.
- `poker_vector1` and `poker_vector2` have different lengths.
- `poker_vector1` and `roulette_vector1` have the same names, while `poker_vector2` and `roulette_vector2` show a names mismatch.
- You can only use [`names()`](http://www.rdocumentation.org/packages/base/functions/names) to _set_ the names of a vector, making `days_vector <- names(poker_vector1)` invalid.

*** =hint
Just try the different code chunks one after the other in the console, and check which statement holds.

*** =pre_exercise_code
```{r}
# NO PEC
```

*** =sct
```{r eval=FALSE}
msg1 <- "The code to define `poker_vector2` is not invalid; try it out in the console!"
msg2 <- "If you use the `length()` function to inspect the length of both `poker_vector1` and `poker_vector2`, you'll see that they both have length 5."
msg3 <- "Correct! You might expect that the names of the vectors `roulette_vector1` and `roulette_vector2` are named the same; but the different approaches treat missing name information differently. Also, notice here how you can also use `names()` to <i>get</i> the names of a vector!"
msg4 <- "Incorrect. You can use `names()` both to <i>set</i> as to <i>get</i> the names of a vector."
test_mc(3, feedback_msgs = c(msg1, msg2, msg3, msg4))
```


--- type:NormalExercise xp:100 skills:1
## Summing and subtracting vectors

Now that you have the poker and roulette winnings nicely as a named vector, you can start doing some data science magic. 

You want to find out the following type of information:
- How much has been your overall profit or loss per day of the week?
- Have you lost money over the week in total?
- Are you winning/losing money on poker or on roulette?

You'll have to do arithmetic calculations on vectors to solve these problems. Remember that this happens element-wise; the following three statements are completely equivalent:

```
c(1, 2, 3) + c(4, 5, 6)
c(1 + 4, 2 + 5, 3 + 6)
c(5, 7, 9)
```

*** =instructions
- Take the element-wise sum of the variables `A_vector` and `B_vector` and it assign to `total_vector`. The result should be a vector.
- Inspect the result by printing `total_vector` to the console.
- Do the same thing, but this time subtract `B_vector` from `A_vector` and assign the result to `diff_vector`.
- Finally, print `diff_vector` to the console as well.

*** =hint
Use the `+` operator to sum `A_vector` and `B_vector`. Use `<-` to assign the result to `total_vector`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# A_vector and B_vector have already been defined for you
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector: total_vector

  
# Print total_vector


# Calculate the difference between A_vector and B_vector: diff_vector


# Print diff_vector

```

*** =solution
```{r eval=FALSE}
# A_vector and B_vector have already been defined for you
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector: total_vector
total_vector <- A_vector + B_vector
  
# Print total_vector
total_vector

# Calculate the difference between A_vector and B_vector: diff_vector
diff_vector <- A_vector - B_vector

# Print diff_vector
diff_vector
```

*** =sct
```{r eval=FALSE}
msg <- "Do not remove or change the definition of the vectors `A_vector` and `B_vector`. These have already been defined for you."
test_object("A_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("B_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("total_vector")
test_output_contains("total_vector", incorrect_msg = "Don't forget to print `total_vector` to the console.")
test_object("diff_vector")
test_output_contains("total_vector", incorrect_msg = "Don't forget to print `diff_vector` to the console.")
success_msg("Good job! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1
## Selection by index (3) 

Now, selecting multiple successive elements of `poker_vector` with `c(2,3,4)` is not very convenient. Many statisticians are lazy people by nature, so they created an easier way to do this: `c(2,3,4)` can be abbreviated to `2:4`, which generates a vector with all natural numbers from 2 up to 4. Try it out in the console!

So, another way to find the mid-week results is `poker_vector[2:4]`. Notice how the vector `2:4` is placed between the square brackets to select element 2 up to 4. You don't have to use the [`c()`](http://www.rdocumentation.org/packages/base/functions/c) function if you're using the shortcut with the colon.

*** =instructions
- Assign to `roulette_subset` the roulette results from Tuesday to Friday inclusive by making use of `:`.
- Print the resulting variable to the console.

*** =hint
Assign a selection of `roulette_vector` to `roulette_subset` by placing `2:5` between square brackets.

*** =sample_code

```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Roulette results for Tuesday to Friday inclusive: roulette_subset


# Print roulette_subset

```

*** =solution

```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Roulette results for Tuesday to Friday inclusive: roulette_subset
roulette_subset <- roulette_vector[2:5]

# Print roulette_subset
roulette_subset
```

*** =sct
```{r eval=FALSE}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_subset",
            incorrect_msg = "It looks like `roulette_selection_vector` does not contain the correct selection from `roulette_vector`. Make sure to to use the right indices.")
test_student_typed("2:5", not_typed_msg = "Make sure to use the combination `2:5` inside square brackets to select the elements in a concise way.")
test_output_contains("roulette_subset", incorrect_msg = "Don't forget to print out `roulette_subset` to the console.")
success_msg("Awesome! The colon operator is extremely useful and very often used in R programming, so remember it well. Have you noticed that the elements in `poker_vector` and `roulette_vector` also have names associated with them? You can also subset vectors using these names, remember?")
```

--- type:NormalExercise xp:100 skills:1
## Selection by name (2)

Just like selecting single elements using numerics extends naturally to selecting multiple elements, you can also use a vector of names. As an example, try

```
roulette_vector[c("Monday","Wednesday")]
```

Of course you can't use the colon trick here: `"Monday":"Wednesday"` will generate an error.

*** =instructions
- Create a vector containing the poker gains for the first three days of the week; name it `poker_start`.
- Using the function [`mean()`](http://www.rdocumentation.org/packages/base/functions/mean), calculate the average poker gains during these first three days. Assign the result to a variable `avg_poker_start`.

*** =hint
You can use `mean(my_vector)` to get the mean of the vector `my_vector`.

*** =pre_exercise_code
```{r eval=FALSE}
# no pec
```

*** =sample_code
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select the first three elements from poker_vector: poker_start


# Calculate the average poker gains during the first three days: avg_poker_start

```

*** =solution
```{r eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select the first three elements from poker_vector: poker_start
poker_start <- poker_vector[c("Monday", "Tuesday", "Wednesday")]

# Calculate the average poker gains during the first three days: avg_poker_start
avg_poker_start <- mean(poker_start)
```

*** =sct
```{r eval=FALSE}

msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("poker_start")
test_object("avg_poker_start",
            incorrect_msg = "Check your code to calculate `avg_poker_start` again. Use the function `mean()` in combination with `poker_start`")
success_msg("Good job! Next to subsetting vectors by index or by name, you can also use logical vectors. The next exercises will test you on this.");
```

--- type:NormalExercise xp:100 skills:1
## Selection by logicals (3)

In this exercise, you'll do a final analysis of your casino ventures. This time, you'll use your newly acquired skills to perform advanced selection on `roulette_vector`. 

Along the way, you'll need the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function. You used it before to calculate the total winnings, so an a numeric vector. However, you can also use [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) on a logical vector; it simply counts the number of vector elements that are `TRUE`.

*** =instructions
- Assign the amounts that you made on the days that you ended positively for roulette to the variable `roulette_profits`. This vector thus contains the positive winnings of `roulette_vector`. You can do this with a one-liner!
- Calculate the sum of the amounts on these profitable days; assign the result to `roulette_total_profit`.
- Find out how many roulette days were profitable, using the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function. Store the result in a variable `num_profitable_days`.

*** =hint
For the first instruction, you can use the exact same technique as in the previous exercise. However, there's no need to specify a selection vector first; you can directly place the comparison inside the square brackets!

*** =pre_exercise_code
```{r, eval=FALSE}
# no pec
```

*** =sample_code
```{r, eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select amounts for profitable roulette days: roulette_profits


# Sum of the profitable roulette days: roulette_total_profit


# Number of profitable roulette days: num_profitable_days

```

*** =solution
```{r, eval=FALSE}
# Casino winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select amounts for profitable roulette days: roulette_profits
roulette_profits <- roulette_vector[roulette_vector > 0]

# Sum of the profitable roulette days: roulette_total_profit
roulette_total_profit <- sum(roulette_profits)

# Number of profitable roulette days: num_profitable_days
num_profitable_days <- sum(roulette_vector > 0)
```

*** =sct
```{r, eval=FALSE}

msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_profits")
test_object("roulette_total_profit")
test_object("num_profitable_days")
success_msg("Awesome! If you inspect the variable num_profitable_days, you'll see that is equal to 2, meaning that you only had two profitable roulette days. You can conclude that roulette is not your game, right?")
```


