--- type:NormalExercise lang:r xp:100 skills:1
## Summarizing a factor

After finishing this course, one of your favourite functions in R will be [`summary()`](http://www.rdocumentation.org/packages/base/functions/summary). This will give you a quick overview of `some_variable`: 

```
summary(some_variable)
``` 

Going back to our survey, you would like to know how many `"Right"` responses you have in your study, and how many `"Left"` responses. The [`summary()`](http://www.rdocumentation.org/packages/base/functions/summary) function gives you the answer to this question.

*** =instructions
Ask a [`summary()`](http://www.rdocumentation.org/packages/base/functions/summary) of `survey_vector` and `survey_factor`. Interpret the results of both vectors. Are they both equally useful in this case?

*** =hint
Call the [`summary()`](http://www.rdocumentation.org/packages/base/functions/summary) function on both `survey_vector` and `survey_factor`, it's as simple as that!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Defintion of survey_vector and survey_factor
survey_vector <- c("R", "L", "L", "R", "R")
survey_factor <- factor(survey_vector, levels = c("R", "L"), labels = c("Right", "Left"))

# Summarize survey_vector


# Summarize survey_factor

```

*** =solution
```{r}
# Defintion of survey_vector and survey_factor
survey_vector <- c("R", "L", "L", "R", "R")
survey_factor <- factor(survey_vector, levels = c("R", "L"), labels = c("Right", "Left"))

# Summarize survey_vector
summary(survey_vector)

# Summarize survey_factor
summary(survey_factor)
```

*** =sct
```{r}
msg = "Do not change anything about the first four lines that define `survey_vector` and `survey_factor`."
test_object("survey_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("survey_factor", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_output_contains("summary(survey_vector)", incorrect_msg = "Make sure to summarize `survey_vector`")
test_output_contains("summary(survey_factor)", incorrect_msg = "Make sure to summarize `survey_factor`")
success_msg("Nice! Have a look at the output. The fact that you identified `\"Right\"` and `\"Left\"` as factor levels in `survey_factor` enables R to show the number of elements for each category.")
```
