--- type:NormalExercise lang:r xp:100 skills:1
## Adding a column for the Worldwide box office 

In the previous exercise you calculated the vector that contained the worldwide box office receipt for each of the three Star Wars movies. However, this vector is not yet part of `star_wars_matrix`.

You can add a column or multiple columns to a matrix with the [`cbind()`](http://www.rdocumentation.org/packages/base/functions/cbind) function, which merges matrices and/or vectors together by column. For example: 

```
big_matrix <- cbind(matrix1, matrix2, vector1 ...)
```

*** =instructions
Add `worldwide_vector` as a new column at the end of the `star_wars_matrix` matrix and assign the result to `star_wars_ext`. Use the [`cbind()`](http://www.rdocumentation.org/packages/base/functions/cbind) function.

*** =hint
Bind the `worldwide_vector` to the `star_wars_matrix` with the `cbind()` function:
```
cbind(the_correct_matrix, the_correct_vector)
```

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Star Wars box office in millions (!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
star_wars_matrix <- rbind(new_hope, empire_strikes, return_jedi)
rownames(star_wars_matrix) <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
colnames(star_wars_matrix) <- c("US", "non-US")

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix: star_wars_ext

```

*** =solution
```{r}
# Star Wars box office in millions (!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
star_wars_matrix <- rbind(new_hope, empire_strikes, return_jedi)
rownames(star_wars_matrix) <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
colnames(star_wars_matrix) <- c("US", "non-US")

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix: star_wars_ext
star_wars_ext <- cbind(star_wars_matrix, worldwide_vector)
```

*** =sct
```{r}

msg = "Do not change anything about the construction and naming of `star_wars_matrix`!"
test_object("star_wars_matrix", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
msg <- "The calculation of `world_wide_vector` has already been coded for you; don't change it."
test_object("worldwide_vector", undefined_msg = msg, incorrect_msg = msg)

test_object("star_wars_ext",
            incorrect_msg = "Did you use the `cbind()` function with the correct arguments to build `star_wars_ext`?")

success_msg("Nice job! After adding column to a matrix, the logical next step is adding rows. Learn how in the next exercise.");
```



--- type:MultipleChoiceExercise lang:r xp:50 skills:1
## Valid syntax or not?

As Filip mentioned in the video, there are different ways to name a matrix. Which one of the following code chunks actually lead to the correct matrix with the correct naming as in the previous exercise? `new_hope`, `empire_strikes` and `return_jedi` are available in the workspace so you can experiment with it.

```
# option A
star_wars_matrix <- rbind(new_hope, empire_strikes, return_jedi)
rownames(star_wars_matrix) <- c("US", "non-US")
colnames(star_wars_matrix) <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# option B
col <- c("US", "non-US")
row <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
rbind(new_hope, empire_strikes, return_jedi, names = c(col, row))

# option C
col <- c("US", "non-US")
row <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), 
                           byrow = TRUE, nrow = 3, dimnames = list(col, row))
                           
# option D
col <- c("US", "non-US")
row <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), 
                           byrow = TRUE, nrow = 3, dimnames = list(row, col))
```

*** =instructions
- Only A correctly creates and names the matrix.
- Both A and C do the trick.
- B and D are the chosen ones.
- Only D is valid syntax that leads to the correct result.

*** =hint
When in doubt, you can always try out the different statements in the console on the right and see which ones work.

*** =pre_exercise_code
```{r}
# Star Wars box office in millions (!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
```

*** =sct
```{r}
msg1 <- "Option A has mixed up row names and column names, so this one won't work. Try again."
msg2 <- "Both option A and C have switched rows and column, so these generate an error. Give it another go."
msg3 <- "Option B is completely invalid syntax. `rbind()` has no `names` argument!"
msg4 <- "Great! Using `dimnames` inside `matrix()` you can construct and name `star_wars_matrix` at the same time."
test_mc(4, feedback_msgs = c(msg1, msg2, msg3, msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1
## Be selective

The workspace already contains the matrix `all_wars_matrix`, containing the box office figures for the six Star Wars movies. It also contains the total revenue for each movie, regardless of the geographical region. 

Have a look at the following R commands:

```
# option A
all_wars_matrix[seq(2, 6, by = 2), "total"]

# option B
all_wars_matrix[c(F,T,F,T,F,T), c(F,T)]

# option C
all_wars_matrix[c("The Empire Strikes Back", 4, 6), c(T,T,F)]

# option D
all_wars_matrix[c(F,T), "total"]
```

Which one of these calls selects the total revenue for the second, fourth and sixth movie in the matrix?

*** =instructions
- A and B
- All four work fine
- A and D
- C and D
- Only B

*** =hint
Just try out the different calls and see which ones give you the selection that is asked for! `all_wars_matrix` is already available in the workspace.

*** =pre_exercise_code
```{r}
# Star Wars box office in millions (!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
star_wars_matrix <- rbind(new_hope, empire_strikes, return_jedi)
colnames(star_wars_matrix) <- c("US", "non-US")
rownames(star_wars_matrix) <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Construct matrix2
box_office_all2 <- c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5)
movie_names2 <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")
star_wars_matrix2 <- matrix(box_office_all2, nrow=3, byrow = TRUE)
colnames(star_wars_matrix2) <- c("US", "non-US")
rownames(star_wars_matrix2) <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
all_wars_matrix <- cbind(all_wars_matrix, total = rowSums(all_wars_matrix))
rm(star_wars_matrix, star_wars_matrix2)
```

*** =sct
```{r}
nok_msg <- "Incorrect. `all_wars_matrix` is available for you to play with, so simply try out the different statements."
ok_msg <- "Great! Notice form code chunk C that it is not possible to combine names and indices in the same vector to perform selection. Up to the next video tutorial!"
test_mc(3, feedback_msgs = c(nok_msg, nok_msg, ok_msg, nok_msg, nok_msg))
```
