--- type:MultipleChoiceExercise xp:50 skills:1
## Vector Subsetting vs. List Subsetting

All these single and double square brackets to perform subsetting on vectors, matrices, factors and now also lists might lead to some confusion. As a summarizing exercise on vector subsetting, consider the following 4 commands.

`shining_list` has been extended with even more information (source: imagination); the list is available in the workspace. Which of the following statements are correct?

 A) `shining_list$box[1,2]` gives the non-US box office of the first release.  
 B) `shining_list[[c(2,4)]]` creates a list containing the actors vector and the box office matrix.  
 C) `shining_list[[c(2,4)]]` returns "Scatman Crothers".  
 D) `shining_list$reviews[1] > shining_list[[3]][3]` is invalid syntax.

*** =instructions
- A and B are correct
- A and C are correct
- A and D are correct
- C and D are correct
- B and D are correct

*** =hint
The dollar sign notation does not need the full name of the element you want to select. `box` is sufficient to refer to the `boxoffice` element of `shining_list`.

*** =pre_exercise_code
```{r, eval=FALSE}
actors_vector <- c("Jack Nicholson","Shelley Duvall","Danny Lloyd","Scatman Crothers","Barry Nelson")
reviews_factor <- factor(c("Good", "OK", "Good", "Perfect", "Bad", "Perfect", "Good"), 
                  ordered = TRUE, levels = c("Bad", "OK", "Good", "Perfect"))
revenue <- matrix(c(39, 47, 18, 14), nrow = 2, byrow = TRUE, dimnames = list(c("First release", "Director's cut"), c("US", "Non-US")))
shining_list <- list(title = "The Shining", actors = actors_vector, reviews = reviews_factor, boxoffice = revenue)
rm(actors_vector, reviews_factor, revenue)
```

*** =sct
```{r, eval=FALSE}
msg1 <- "Wrong. Statement B is false, as you'd need single brackets to subset a list to end up with a new list. Also, remember that `shining_list[[c(2,4)]]` is equivalent to `shining_list[[2]][[4]]`."
msg2 <- "Correct! Next to subsetting lists, you'll also often want to extend lists. Get some hands-on experience in the following exercises."
msg3 <- "Incorrect. The code in statement D is perfectly valid, as you are selecting two ordered factors."
msg4 <- msg3
msg5 <- "Incorrect. Both B and D are false."
test_mc(2, feedback_msgs = c(msg1, msg2, msg3, msg4, msg5))
```


--- type:NormalExercise xp:100 skills:1
## Extending a list (2)

To extend a list with multiple elements at the same time, you can again use the combine function. To add both my opinion and your opinion on the movie:

```
c(shining_list, 
  my_opinion = "Love it!", 
  your_opinion = "Hate it!")
```

This works fine, but what if you want to add an entire vector as an element to the list:

```
c(shining_list,
  opinions = c("Love it!", "Hate it!"))
```

You'll see that a different list element per vector element was created. To fix this, you better surround the elements you want to add to the list in another `list()` function. That way, the combine function will simply paste together two lists (try it out in the console!):

```
c(shining_list, 
  list(my_opinion = "Love it!",
       your_opinion = "Hate it!"))
                     
c(shining_list, 
  list(c("Love it!", "Hate it"!)))
```

*** =instructions
- Add both the year (1980) and director (Stanley Kubrick) to `shining_list` again. Try to do it in a one-liner this time, and save the result in a new variable, `shining_list_ext`.
- Again, have a look at the structure of `shining_list_ext`.

*** =hint
Have a look at the example code in the exercise assignment. Maybe this can help you start:
```
shining_list <- c(shining_list, ...)
```
You still have to add some code where the three dots are.

*** =pre_exercise_code
```{r, eval = FALSE}
# Create shining_list
actors_vector <- c("Jack Nicholson","Shelley Duvall","Danny Lloyd","Scatman Crothers","Barry Nelson")
reviews_factor <- factor(c("Good", "OK", "Good", "Perfect", "Bad", "Perfect", "Good"), 
                  ordered = TRUE, levels = c("Bad", "OK", "Good", "Perfect"))
shining_list <- list(title = "The Shining", actors = actors_vector, reviews = reviews_factor)
rm(actors_vector, reviews_factor)
```

*** =sample_code

```{r eval=FALSE}
# shining_list is already defined in the workspace

# Add both the year and director to shining_list: shining_list_ext


# Have a look at the structure of shining_list_ext

```

*** =solution

```{r eval=FALSE}
# shining_list is already defined in the workspace

# Add both the year and director to shining_list: shining_list_ext
shining_list_ext <- c(shining_list, year = 1980, director = "Stanley Kubrick")
shining_list_ext <- c(shining_list, list(year = 1980, director = "Stanley Kubrick"))  # also works

# Have a look at the structure of shining_list_ext
str(shining_list_ext)
```

*** =sct
```{r eval=FALSE}
test_error()
msg = "Do not remove or override `shining_list`."
test_object("shining_list", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("shining_list_ext",
            incorrect_msg = "It looks like `shining_list_ext` does not contain all the correct elements. It should contain both the year (1980) as well as the director (Stanley Kubrick)! Make sure to adopt the same order.")
test_object("shining_list_ext", eq_condition = "equal",
            incorrect_msg = "It looks like `shining_list_ext` does not contain the correct naming for all the components.");

test_output_contains("str(shining_list_ext)", incorrect_msg = "Make sure to inspect the structure of `shining_list_ext`.")
success_msg("Great!")
```
