--- type:MultipleChoiceExercise lang:r xp:50 skills:1
## Valid syntax or not?

As Filip mentioned in the video, there are different ways to name a matrix. Which one of the following code chunks actually lead to the correct matrix with the correct naming as in the previous exercise? `new_hope`, `empire_strikes` and `return_jedi` are available in the workspace so you can experiment with it.

```
# option A
star_wars_matrix <- rbind(new_hope, empire_strikes, return_jedi)
rownames(star_wars_matrix) <- c("US", "non-US")
colnames(star_wars_matrix) <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# option B
col <- c("US", "non-US")
row <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
rbind(new_hope, empire_strikes, return_jedi, names = c(col, row))

# option C
col <- c("US", "non-US")
row <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), 
                           byrow = TRUE, nrow = 3, dimnames = list(col, row))
                           
# option D
col <- c("US", "non-US")
row <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), 
                           byrow = TRUE, nrow = 3, dimnames = list(row, col))
```

*** =instructions
- Only A correctly creates and names the matrix.
- Both A and C do the trick.
- B and D are the chosen ones.
- Only D is valid syntax that leads to the correct result.

*** =hint
When in doubt, you can always try out the different statements in the console on the right and see which ones work.

*** =pre_exercise_code
```{r}
# Star Wars box office in millions (!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
```

*** =sct
```{r}
msg1 <- "Option A has mixed up row names and column names, so this one won't work. Try again."
msg2 <- "Both option A and C have switched rows and column, so these generate an error. Give it another go."
msg3 <- "Option B is completely invalid syntax. `rbind()` has no `names` argument!"
msg4 <- "Great! Using `dimnames` inside `matrix()` you can construct and name `star_wars_matrix` at the same time."
test_mc(4, feedback_msgs = c(msg1, msg2, msg3, msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1
## Be selective

The workspace already contains the matrix `all_wars_matrix`, containing the box office figures for the six Star Wars movies. It also contains the total revenue for each movie, regardless of the geographical region. 

Have a look at the following R commands:

```
# option A
all_wars_matrix[seq(2, 6, by = 2), "total"]

# option B
all_wars_matrix[c(F,T,F,T,F,T), c(F,T)]

# option C
all_wars_matrix[c("The Empire Strikes Back", 4, 6), c(T,T,F)]

# option D
all_wars_matrix[c(F,T), "total"]
```

Which one of these calls selects the total revenue for the second, fourth and sixth movie in the matrix?

*** =instructions
- A and B
- All four work fine
- A and D
- C and D
- Only B

*** =hint
Just try out the different calls and see which ones give you the selection that is asked for! `all_wars_matrix` is already available in the workspace.

*** =pre_exercise_code
```{r}
# Star Wars box office in millions (!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
star_wars_matrix <- rbind(new_hope, empire_strikes, return_jedi)
colnames(star_wars_matrix) <- c("US", "non-US")
rownames(star_wars_matrix) <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Construct matrix2
box_office_all2 <- c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5)
movie_names2 <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")
star_wars_matrix2 <- matrix(box_office_all2, nrow=3, byrow = TRUE)
colnames(star_wars_matrix2) <- c("US", "non-US")
rownames(star_wars_matrix2) <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
all_wars_matrix <- cbind(all_wars_matrix, total = rowSums(all_wars_matrix))
rm(star_wars_matrix, star_wars_matrix2)
```

*** =sct
```{r}
nok_msg <- "Incorrect. `all_wars_matrix` is available for you to play with, so simply try out the different statements."
ok_msg <- "Great! Notice form code chunk C that it is not possible to combine names and indices in the same vector to perform selection. Up to the next video tutorial!"
test_mc(3, feedback_msgs = c(nok_msg, nok_msg, ok_msg, nok_msg, nok_msg))
```
